### 2016-09-19 reports/agronomic_data.Rmd
Do not display code chunks in report

# Agronomic Data
```{r echo = FALSE, message = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "gridExtra", "modelr", "caret",
               "knitr", "tibble")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
                      warning = FALSE, echo = FALSE)
```

```{r echo = FALSE}
# Load data
pheno <- readRDS("./data/processed/Pheno_stage2.RDS")

# Prepare agronomic data.
genos <- readRDS("./data/processed/common_genotypes.RDS")
hybrid <- genos$Hybrid
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% hybrid) %>%
  select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))

# Select hybrids for which both parents have genomic as well as transcriptomic
# records.
dent <- genos$Dent$mrna
flint <- genos$Flint$mrna
hybrid <- genos$Hybrid %>%
  as_data_frame() %>%
  separate(value, into = c("DF", "Dent", "Flint")) %>%
  mutate(Avail_Dent = ifelse(Dent %in% dent, yes = 1, no = 0),
         Avail_Flint = ifelse(Flint %in% flint, yes = 1, no = 0),
         Tested_Parents = Avail_Dent + Avail_Flint) %>%
  filter(Tested_Parents == 2) %>%
  unite(Hybrid, DF, Dent, Flint) %>%
  .$Hybrid

# All hybrids
all_pheno <- all_pheno %>%
  mutate(Reduced_Data = "no")

# Only hybrids for which parents have both, genomic and transcriptomic data
reduced_pheno <- all_pheno %>%
  mutate(Reduced_Data = ifelse(G %in% hybrid, yes = "yes", no = "no")) %>%
  filter(Reduced_Data == "yes")
```

## Skewness
A general **rule of thumb** to consider is that skewed data whose ratio of the 
highest value to the lowest value is greater than 20 have significant skewness.
The formula for the skewness statistic is
$$
\text{skewness}  = \frac{\sum (x_i - \bar{x})^{3}}{(n - 1)v^{3/2}},
$$
where $v = \frac{\sum (x_i - \bar{x})^2}{(n - 1)}$ with $x$ is the predictor
variable, $n$ is the number of values and $\bar{x}$ is the sample mean of the 
predictor.
```{r Agro-Skewness, echo = FALSE}
all_skewness <- all_pheno %>%
  group_by(Trait) %>%
  summarize(Skewness = e1071::skewness(Observed),
            Skewness = round(Skewness, digits = 2))
reduced_skewness <- reduced_pheno %>%
  group_by(Trait) %>%
  summarize(Skewness = e1071::skewness(Observed),
            Skewness = round(Skewness, digits = 2))
full_join(x = all_skewness, y = reduced_skewness,
          by = "Trait",
          suffix = c(" (Full)", " (Reduced)")) %>%
  kable(padding = 0)
```

## Distribution
```{r Agro-Distribution, message = FALSE, echo = FALSE, fig.cap = "Histogram of the distribution of seven agronomic traits. Black bins show the distribution of the reduced set of hybrids whereas light grey bins show the distribtution of the full set of hybrids."}
ggplot(all_pheno, aes(x = Observed)) +
  geom_histogram(fill = "grey", alpha = 0.5) +
  geom_histogram(data = reduced_pheno, color = "black") +
  facet_wrap(~ Trait, scales = "free") +
  theme_bw()
```

## Correlation
```{r Agro-Correlation, echo = FALSE, fig.cap = "Pairwise correlations among seven agronomic traits for the full set of hybrids and the reduced set of hybrids, respectively."}
par(mar = c(0, 1, 0, 0),
    mfrow = c(1, 2))
all_pheno %>%
  select(-Reduced_Data) %>%
  spread(key = Trait, value = Observed) %>%
  as.data.frame() %>%
  remove_rownames() %>%
  column_to_rownames(var = "G") %>%
  cor() %>%
  corrplot::corrplot(col = viridis(256),
                     method = "color",
                     diag = FALSE,
                     cl.pos = "b",
                     tl.pos = "d",
                     tl.srt = 60,
                     order = "hclust",
                     addCoef.col = "black",
                     title = "Full",
                     mar = c(0, 1, 1, 0))
reduced_pheno %>%
  filter(Reduced_Data == "yes") %>%
  select(-Reduced_Data) %>%
  spread(key = Trait, value = Observed) %>%
  as.data.frame() %>%
  remove_rownames() %>%
  column_to_rownames(var = "G") %>%
  cor() %>%
  corrplot::corrplot(col = viridis(256),
                     method = "color",
                     diag = FALSE,
                     cl.pos = "b",
                     tl.pos = "d",
                     tl.srt = 60,
                     order = "hclust",
                     addCoef.col = "black",
                     title = "Reduced",
                     mar = c(0, 1, 1, 0))
```