### 2016-09-20 reports/conclusions.Rmd
Update references in report

# Conclusions and Outlook
We have seen that

1.    for the 'reduced' set of hybrids, predictive abilities for the traits 
      'DMY', 'PRO' and 'FAT' were higher when predicted via mRNA compared to SNP
      data (Table \@ref(tab:Non-Imputed-Pred-Ability)).
2.    for the 'full' set of hybrids, predictive abilities based on single-step
      prediction with transcriptomic data still outperformed SNP data for the
      traits 'DMY' and 'PRO' (Table \@ref(tab:Imputed-Pred-Ability)).
2.    an exploration of the effect of different levels of missingness &mdash; 
      with respect to transcriptomic BLUEs &mdash; in the parents is 
      computationally prohibitive 
      (Chapter \@ref(prediction-with-imputed-mrnas)).

Currently, our primary goal is to demonstrate the utility of imputing missing 
data in a quantitative predictor (exemplified by transcriptomic data) and 
using the resulting feature matrix in single-step hybrid prediction.