### 2016-09-20 reports/data_overview.Rmd

# Data Overview
```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "knitr", "tibble", "purrr")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

```{r Load-Agro-Data, message = FALSE, cache = TRUE}
pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
genos <- readRDS("./data/processed/common_genotypes.RDS")

full_hybrid <- genos$Hybrid
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% full_hybrid) %>%
  dplyr::select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))
traits <- all_pheno %>%
  .$Trait %>%
  levels()

# Hybrids of parents with records for both endophenotypes
red_dent <- Reduce(intersect, genos[["Dent"]][c("snp", "mrna")])
red_flint <- Reduce(intersect, genos[["Flint"]][c("snp", "mrna")])
red_hybrid <- genos$Hybrid %>%
  as_data_frame() %>%
  separate(value, into = c("DF", "Dent", "Flint")) %>%
  mutate(Avail_Dent = ifelse(Dent %in% red_dent, yes = 1, no = 0),
         Avail_Flint = ifelse(Flint %in% red_flint, yes = 1, no = 0),
         Tested_Parents = Avail_Dent + Avail_Flint) %>%
  filter(Tested_Parents == 2) %>%
  unite(Hybrid, DF, Dent, Flint) %>%
  .$Hybrid
full_hybrid <- genos$Hybrid
genos$Hybrid <- NULL
genos$Hybrid$snp <- full_hybrid
genos$Hybrid$mrna <- red_hybrid
```

```{r Feature-Number, echo = FALSE}
snp <- readRDS("./data/processed/snp_mat.RDS")
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
```

Agronomic data for `r length(traits)` traits (including dry-matter yield (DMY))
and `r length(full_hybrid)` hybrids are available 
(Table \@ref(tab:Geno-Number)).
After applying a series of quality checks, `r ncol(snp)` SNPs and 
best linear unbiased estimates of `r ncol(mrna)` were available for further 
analyses.
Both, genomic and transcriptomic records, were collected for inbred parent lines
of `r length(red_hybrid)` hybrids.

```{r Geno-Number, echo = FALSE}
geno_tab_cap <- paste("Number of genotypes for Dent and Flint parent lines and",
                      "their hybrid progeny for which genomic and/or",
                      "transcriptomic data are available.",
                      "All genotypes for which transcriptomic data are", 
                      "available, also have genomic data.",
                      "Agronomic data are available for all 1,521 hybrid",
                      "genotypes.")
# Number of genotypes for either SNPs or mRNAs separated by hybrids and their
# parental Dent and Flint lines, respectively.
genos %>% 
  at_depth(.depth = 2, .f = length) %>%
  bind_rows() %>%
  dplyr::select(-agro) %>%
  mutate(Type = names(genos)) %>%
  dplyr::select(Type, snp, mrna) %>%
  rename("Genomic Data" = snp,
         "Transcriptomic Data" = mrna) %>%
  knitr::kable(caption = geno_tab_cap, booktabs = TRUE)
```