### 2016-09-21 reports/appendix.Rmd
Init GitBook

```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("knitr")
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

# Additional Information
## Gamazon et al. (Nature Genetics, 2015)
## Rationale
*    Proposition: directly test the molecular mechanisms through which genomic 
     information affect phenotypic traits.
*    Gene regulation information is missing for some genotypes whereas genomic 
     information are complete.
     The latter was used to to impute missing gene expression information.
*    A major advantage of this approach is that less (imputed) features can be 
     used as surrogates for genomic information in phenotype prediction, thereby
     alleviating the multiple testing problem.
*    Estimate genetically regulated gene expression (GReX), which constitute 
     multiple-SNP prediction of expression levels.
*    Use GReX to impute missing transcriptomic data.

### Shortcomings
"We are aware of the attenuation bias that arises because of the error in the 
estimation of GRex.
This is a subject to be investigated in the future, but the bias does not 
invalidate our analysis as we only use the estimate of GReX as a discovery 
tool."

## Method
The additive model of gene expression is
$$
T = \sum_{k=1}^{K} w_{k} X_{k} + \epsilon,
$$
where $T$ denotes the vector of gene expression values, $k$ denotes the SNP
marker, $X_{k}$ is the number of reference alleles for $k$ and $w_{k}$ is the
coefficient of the model to be fitted.
$\mathbf{w}$ was estimated using LASSO and elastic net.

Missing gene expression values (GReX) are then imputed by multiplying the matrix 
of reference alleles for SNPs with the coefficients $\mathbf{w}$ from the 
training set.
Finally, GReX and the trait of interested are correlated.

## Garrick, Taylor and Fernando (2009)
1.    We do not have true genetic values at our disposal.
2.    The approach of @Gamazon2015 does not account for the non-diagonal 
      variance-covariance structure of the residuals.

Issues with BLUP:
1.    Superior genotypes tend to be underevaluated (i.e. have negative 
      prediction errors) whereas inferior animals tend to be overevaluated.
2.    Observations are shrunken towards the mean while the extent of shrinkage
      depends upon the amount of information, i.e. the reliability.
3.    Training genotypes, which were selected according to their estimated 
      breeding values, comprise individuals with varying reliability.