### 2016-09-26 reports/impute_single_mrna.Rmd
Init mRNA report

# Genotype-Wise mRNA-Imputation

## Function for the imputation of a single mRNA.
```{r Impute-mRNA-Function}
## Function for the imputation of a single mRNA
impute_one_mrna <- function(x, y) {
  geno <- rownames(x)
  # Names of genotypes for which transcriptomic records exist
  nm2 <- rownames(y)
  # Names of genotypes for which transcriptomic records are missing.
  nm1 <- setdiff(rownames(x), nm2)
  # Hybrid parents not in y.
  geno1 <- geno[geno %in% nm1]
  # Hybrid parents in y.
  geno2 <- geno[geno %in% nm2]
  x <- x[match(unique(geno), rownames(x)), ]
  # Design matrix mapping GCA effects from genotypes with transcriptomic
  # records to y2.
  Z2 <- Matrix::sparse.model.matrix(~-1 + factor(geno2),
                                    drop.unused.levels = FALSE)
  colnames(Z2) <- gsub("factor\\(geno2\\)", replacement = "", x = colnames(Z2))
  Z2 <- Z2[, match(nm2, colnames(Z2))]
  rownames(Z2) <- geno2
  # Hybrid parents not in y.
  geno1 <- geno[geno %in% nm1]
  # Hybrid parents in y.
  geno2 <- geno[geno %in% nm2]
  x <- x[match(unique(geno), rownames(x)), ]
  y <- y[nm2, ]
  M2 <- y[, matrixStats::colVars(y) != 0]
  x <- x[, matrixStats::colVars(x) != 0]
  A <- build_kernel(M = x, lambda = 0.01, algorithm = "RadenII")
  A11 <- A[nm1, nm1, drop = FALSE]
  A12 <- A[nm1, nm2, drop = FALSE]
  A21 <- A[nm2, nm1, drop = FALSE]
  A22 <- A[nm2, nm2, drop = FALSE]
  Ainv <- solve(A)
  dimnames(Ainv) <- dimnames(A)
  A_up11 <- Ainv[nm1, nm1]
  A_up12 <- Ainv[nm1, nm2]
  # Eq.21
  M1 <- A12 %*% solve(A22) %*% M2
  J2 <- matrix(-1, nrow = ncol(A12), ncol = 1)
  # Eq.22
  J1 <- A12 %*% solve(A22) %*% J2
  # Eq.10
  epsilon <- t(chol(solve(Ainv[nm1, nm1])))
  # Eq.20
  W2 <- Z2 %*% M2
  w1 <- A12 %*% solve(A22) %*% Z2 %*% M2
  as.numeric(w1)
}
```

Plot the distribution of correlations between vectors of original and imputed
mRNAs.
```{r Correlation-Imputed-vs-NonImputed}
#imp_mrna <- vapply(seq_len(nrow(y)), FUN = function(i) {
#  z <- y[-i, ]
#  imp_mrna <- impute_one_mrna(x, z)
#  cor(y[i, ], imp_mrna)
#}, FUN.VALUE = numeric(1))
#hist(imp_mrna)
```