### 2016-09-27 reports/loocv_example.Rmd

# Efficient LOOCV With Prediction Errors

The following code snippet is taken from the [BGLR Vignette](https://github.com/gdlc/BGLR-R/blob/master/inst/md/hyperparameters.md).
```{r Fit-BGLR-Model, cache = TRUE}
pacman::p_load("BGLR")
data("mice")
X <- scale(mice.X[, 1:2000])
h2 <- 0.5
QTL <- seq(from = 50, to = ncol(X), length = 20)
nQTL <- length(QTL)
n <- nrow(X)
b <- rep(1, nQTL) * sqrt(h2 / nQTL)
signal <- X[, QTL] %*% b
error <- rnorm(n, sd = sqrt(1 - h2))
y <- signal + error

## Using a flat prior
fm1 <- BGLR(y = y, ETA = list(list(X = X, model = "BayesB", counts0 = 1,
                                   counts1 = 1)), 
            nIter = 5000, burnIn = 500, verbose = FALSE)
```

All formulae are from Cheng, Garrick and Fernando (2016, unpublished):
"Efficient Strategies for Leave-one-out Cross Validation for Genomic Best Linear
Unbiased Prediction".

$$
\mathbf{X}^{*} = 
  \begin{bmatrix}
    \mathbf{1} & \mathbf{X}
  \end{bmatrix}
$$
```{r X-prime}
X_prime <- cbind(matrix(1, nrow = nrow(X), ncol = 1),
                 X)
```

$$
\hat{\beta}^{*} =
  \begin{bmatrix}
    \hat{\mu} \\
    \hat{\beta}
  \end{bmatrix}
$$
```{r}
b_hat <- fm1$ETA[[1]]$b
mu_hat <- fm1$mu
b_hat_prime <- matrix(c(mu_hat, b_hat), ncol = 1)
```

$\lambda = \frac{\sigma_{e}^{2}}{\sigma_{\beta}^{2}}$
```{r}
lambda <- c(fm1$varE / fm1$ETA[[1]]$varB, 1)
```

Diagonal matrix $\mathbf{D}_{[(p + 1) \times (p + 1)]}$.
```{r}
Diag <- cbind(diag(nrow = ncol(X), ncol = ncol(X)),
              matrix(1, nrow = ncol(X), ncol = 1))
Diag <- rbind(Diag,
              matrix(1, nrow = 1, ncol = ncol(Diag)))
```

$\mathbf{H} = \mathbf{X}^{*}(\mathbf{X}^{*'}\mathbf{X}^{*} +
\mathbf{D}\lambda)^{-1}\mathbf{X}^{*'}$
```{r H, cache = TRUE}
H <- X_prime %*% solve(t(X_prime) %*% X_prime + Diag * lambda) %*% t(X_prime)
```

$\hat{e_{j}} = \frac{y_{j} - 
\mathbf{x^{*'}}_{j}\hat{\beta}^{*}}{1-\mathbf{H_{jj}}}$
```{r LOOCV-Residuals}
e_hat <- vapply(seq_len(nrow(X_prime)), FUN = function(j) {
  t_x_prime <- X_prime[j, , drop = FALSE]
  numerator <- as.numeric(y[j] - t_x_prime %*% b_hat_prime)
  denominator <- 1 - H[j, j]
  numerator / denominator
}, FUN.VALUE = numeric(1))
```

$\text{PRESS} = \sum_{i=1}^{n} \hat{e}_{j}^{2}$
```{r LOOCV-Output-Summary}
press <- sum(e_hat ^ 2)
rmse <- sqrt(press / length(e_hat))
y_hat <- y - e_hat
# Leave-one-out prediction accuracy
r <- cor(y, y_hat)
```

The LOOCV-RMSE is `r round(rmse, digits = 2)` and the prediction accuracy is 
`r round(r, digits = 2)`.