### 2016-10-21 reports/loocv_issues.Rmd

---
title: "Fernando LOOCV"
author: "Matthias Westhues"
date: "`r Sys.Date()`"
output: pdf_document
---

My assumption is that the efficient implementation of LOOCV proposed by Cheng
et al. (2016) uses the entire sample space except for a single hybrid that 
is used as a test set.
However, our goal is to apply LOOCV only to the T0 hybrids.
In this case, not very hybrid, that was previously part of the training set to
predict another hybrid, can be moved to the test set (see the following 
example).

|:---:|:----:|
|ID   |Hybrid|
|1    |Aa    |
|2    |Ab    |
|3    |Ac    |
|4    |Ba    |
|5    |Bb    |
|6    |Bc    |
|7    |Ca    |
|8    |Cb    |
|9    |Cc    |

: Hybrids and their corresponding IDs.

|:--------:|:------------:|
| Test Set | Training Set |
| 1        | 5, 6, 8, 9   |        
| 2        | 4, 6, 8, 9   |        
| 3        | 4, 5, 7, 8   |        
| 4        | 2, 3, 8, 9   |        
| 5        | 1, 3, 7, 9   |        
| 6        | 1, 2, 7, 8   |        
| 7        | 2, 3, 5, 6   |        
| 8        | 1, 3, 4, 6   |        
| 9        | 1, 2, 4, 5   |

: Hybrids in the test set and their corresponding T0-Hybrids in the training 
set. The hybrids are denoted by their IDs from the previous table.

Hybrid \#1 can be predicted with hybrids 5, 6, 8 and 9, however, hybrid \#5 
cannot be predicted with hybrids 6 and 8, hybrid \#6 cannot be predicted with 
hybrids 5, 6 and 9, hybrid \#8 cannot be predicted with hybrids 5 and 9, and 
hybrid \#9 cannot be predicted with hybrids 6 und 8.

## Proposed Solution