### 2016-10-21 reports/transcriptomic_data.Rmd

# Transcriptomic Data
Keep only genotypes for which both, transcriptomic and genomic data are 
available.
```{r echo = FALSE}
pacman::p_load("sspredr", "tidyverse", "caret", "e1071", "viridis")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", small.mark = ".", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
                      warning = FALSE, message = FALSE, echo = FALSE)
```

```{r mRNA-Data, cache = TRUE}
# Common genotypes
genos <- readRDS("./data/processed/common_genotypes.RDS")
mrna_inbreds <- genos %>%
  filter(Pool != "Hybrid", Data_Type == "mrna") %>%
  split(.$Pool) %>%
  map("G")

# mRNA
mrna <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna <- t(mrna)
mrna <- mrna[grep("Exp|Sigma", x = rownames(mrna), invert = TRUE), ]
mrna_df <- mrna %>%
  as_data_frame() %>%
  mutate(G = rownames(mrna)) %>%
  mutate(Group = ifelse(G %in% mrna_inbreds$Dent, yes = "Dent", no = "Flint"),
         Group = as.factor(Group)) %>%
  select(-G)
```

Separately for each heterotic group, run a principal component analysis (PCA) on 
the "raw" mRNA-BLUEs and return the importance of the first three PCs.
```{r Raw-PCA, cache = TRUE}
rm_grp <- function(x) {
  x$Group <- NULL
  x
}
pca_var_imp <- function(x) {
  var_imp <- x$sd ^ 2 / sum(x$sd ^ 2) * 100
  sum(var_imp[1:3])
}
# Variable importance in PCA (first three PCs)
raw_varimp <- mrna_df %>%
  split(.$Group) %>%
  map(rm_grp) %>%
  map(prcomp) %>%
  map(pca_var_imp) %>%
  bind_rows(.id = "Type") %>%
  mutate(Type = "Raw")
```

## Data pre-processing
Pre-process the data in the following order:

1.    Near-zero-variance filtering.
2.    Box-Cox tranformation
3.    Centering
4.    Scaling
```{r Transform-Data, cache = TRUE}
mrna_lst <- mrna_df %>% 
  split(.$Group) %>%
  map(rm_grp) %>%
  map(~as.matrix(.))

# Data transformation
trans_lst <- mrna_lst %>%
  map(preProcess,
      method = c("BoxCox", "center", "scale", "nzv")) %>%
  map2(.y = mrna_lst, .f = predict)

# Variable importance in PCA (first three PCs)
trans_varimp <- trans_lst %>%
  map(prcomp) %>%
  map(pca_var_imp) %>%
  bind_rows(.id = "Type") %>%
  mutate(Type = "Trans")

# Compare variable importances for raw and transformed mRNA data.
rbind(raw_varimp, trans_varimp) %>%
  knitr::kable(caption = "Importance of the first three PCs (in %)",
               digits = 2)
```

```{r Tidy-Data, cache = TRUE}
mrna_lst <- list(Raw = mrna_lst, Trans = trans_lst)
tidy_mrna <- mrna_lst %>% 
  at_depth(.depth = 2, .f = ~as.data.frame(.)) %>%
  at_depth(.depth = 1, .f = ~bind_rows(., .id = "Group")) %>%
  bind_rows(.id = "Type") %>%
  as_tibble() %>%
  gather(key = "mRNA", value = "Expression", -Group, -Type) %>%
  mutate(Type = as.factor(Type),
         Group = as.factor(Group))
```

## Skewness
```{r Check-Normality, cache = TRUE}
mrna_skew <- tidy_mrna %>%
  group_by(Group, Type, mRNA) %>%
  summarize(Skewness = e1071::skewness(Expression),
            Shapiro = stats::shapiro.test(Expression)$statistic)
```

```{r, mRNA-Skewness, cache = TRUE, fig.cap = "Distribution of skewness values across all mRNAs, separately for transformed and raw BLUEs."}
mrna_skew %>%
  ggplot(aes(x = Skewness, fill = Type)) +
  xlim(c(-2, 2)) +
  geom_density(alpha = 0.2) +
  facet_grid(. ~ Group) +
  theme(legend.position = "top") +
  ylab("Density")
```

```{r Association-Shapiro-Wilks-Skewness, cache = TRUE, fig.cap = "Shapiro-Wilks Test Statistic vs absolute skewness for transcriptomic BLUEs depending on whether the BLUEs were transformed or not."}
mrna_skew %>%
  mutate(Abs_Skewness = abs(Skewness)) %>%
  ggplot(aes(x = Shapiro, y = Abs_Skewness)) +
  xlim(c(0.8, 1)) +
  ylim(c(0, 2)) +
  geom_point() +
  facet_grid(Group ~ Type) +
  xlab("Shapiro-Wilks Test Statistic") +
  ylab("Absolute Skewness")
```