### 2016-11-22 reports/predictor_subsetting.Rmd

# Predictor Subsetting
```{r setup-predictor-subsets, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
               "stringr", "forcats")

devtools::install_github("mwesthues/sspredr")
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

Goal: Generate nine subsets of predictors for later comparisons between their
predictive abilities:

a.    100% SNP
b.    100% pedigree
c.    42% mRNA
d.    42% pedigree
e.    42% SNP
f.    100% pedigree & 77% SNP
g.    100% pedigree & 42% mRNA
h.    100% SNP & 42% mRNA
i.    100% pedigree & 77% SNP & 42% mRNA

Make comparisons between the following scenarios:

1.    a, b, f, g, h, i
2.    c, d, e

Throughout this manuscript, digits behind the variable name of each predictor
(ped: pedigree, snp: genomic, mrna: transcriptomic) indicate the size of 
the subset relative to the full set of genotypes (n = 245).
For example, we have transcriptomic data for 103 genotypes, which are roughly 
42% of 245 genotypes, hence the name 'mrna42' and so forth.
For the set "snp77", all genotypes &mdash; for which transcriptomic records are 
available &mdash; were used and the remaining genotypes were sampled at random
with equally large proportions being sampled from the Dent and the Flint 
heterotic group, respectively.

```{r Prepare-Predictor-Subsetting, cache = TRUE, results = "hide"}
# -- Genotype information.
# Information on the set of genotypes.
genos <- readRDS("./data/processed/common_genotypes.RDS")
genos %>%
  filter(Pool %in% c("Dent", "Flint")) %>%
  group_by(Pool, Data_Type) %>%
  count()

# Get the names of the Dent Hybrid-parents.
geno <- genos %>%
  filter(Pool == "Hybrid") %>%
  .$G %>%
  map(~stringr::str_split(., pattern = "_")) %>%
  at_depth(.depth = 2, .f = 2) %>%
  flatten() %>%
  as_vector()

# -- Load predictor data.
# pedigree
ped100 <- readRDS("./data/processed/ped-datafull-GTP.RDS")
ped_nms <- genos %>%
  filter(Data_Type == "ped") %>%
  .$G
ped100 <- as.matrix(ped100) * 2
ped100 <- ped100 %>%
  .[rownames(.) %in% ped_nms, colnames(.) %in% ped_nms]
ped100[is.na(ped100)] <- 0

# snp
snp100 <- readRDS("./data/processed/snp_mat.RDS")
snp_nms <- genos %>%
  filter(Data_Type == "snp") %>%
  .$G
snp100 <- snp100 %>%
  .[rownames(.) %in% snp_nms, ]

# mrna
mrna42 <- readRDS("./data/processed/subset_mrna_blues.RDS")[["100%"]]
mrna_nms <- genos %>%
  filter(Data_Type == "mrna") %>%
  .$G
mrna42 <- mrna42 %>%
  as.matrix() %>%
  t() %>%
  .[rownames(.) %in% mrna_nms, ]
```

```{r Resample-Predictor-Data, cache = TRUE, results = "hide"}
## Resample predictor data 
# Remove some genomic records so that we can impute them via pedigree 
# information.
set.seed(314)
snp35 <- snp100 %>%
  as_tibble() %>%
  mutate(G = rownames(snp100)) %>%
  filter(!G %in% rownames(mrna42)) %>%
  left_join(genos %>% filter(Data_Type == "snp") %>% select(-Data_Type),
            by = "G") %>%
  group_by(Pool) %>%
  sample_frac(size = 0.6) %>%
  ungroup() %>%
  select(-Pool) %>%
  as.data.frame() %>%
  column_to_rownames(var = "G") %>%
  as.matrix()
snp77 <- snp100 %>%
  .[rownames(.) %in% rownames(mrna42), ] %>%
  rbind(snp35)
stopifnot(all(rownames(mrna42) %in% rownames(snp77)))

ped42 <- ped100 %>%
  .[match(rownames(mrna42), rownames(.)), 
    match(rownames(mrna42), colnames(.))]
snp42 <- snp100 %>%
  .[match(rownames(mrna42), rownames(.)), ]

# Store the new objects
storage_dir <- "./data/derived/predictor_subsets/"
ls() %>%
  .[. != "snp35"] %>%
  .[stringr::str_detect(., pattern = "[:digit:]")] %>%
  as.list() %>%
  map_if(.p = ~get(.) %>% is.matrix,
         .f = ~saveRDS(get(.), file = paste0(storage_dir, ., ".RDS"),
                       compress = TRUE))
```