### 2017-02-01 reports/outlook.Rmd
Add outlook on Bustos-Korts_2016

# Outlook
## Size of individual predictors
We have not yet investigated the influence of the size (in terms of the number
of parental genotypes) of an individual predictor on the predictive ability
achieved with a combination of multiple predictors in a single-step prediction
settting.
Consider the following predictors in single-step prediction, where nothing
changes but the coverage of genotypes through genomic data in increments of 10%:

1.   'ped100-snp90' or &nbsp;&nbsp;&nbsp;&nbsp; 'ped100-snp90-mrna42'

2.   'ped100-snp80' or &nbsp;&nbsp;&nbsp;&nbsp; 'ped100-snp80-mrna42'

3.   $\dots$        or  &nbsp;&nbsp;&nbsp;&nbsp; $\dots$

4.   'ped100-snp10' or &nbsp;&nbsp;&nbsp;&nbsp; 'ped100-snp10-mrna42'

Using such an iterative approach, we might get closer to describing the
contribution of individual predictors to the predictive ability in a combined
single-step predictor.

One caveat here is that, so far, we have sampled 'snp77' at random (except for
the adjustment to sampling, in relative terms, equally large fractions from the
Dent and the Flint group, respectively).
If we run the scenario proposed above, we would have a high computational load.
What I am looking for is a way to sample 'snpXX' --- where XX reprents the
fraction of all possible genotypes covered --- such that it covers the genetic
target space (i.e. the whole set of 142 parental Dent and 103 parental Flint
lines) equally well.
@Bustos-Korts2016 have described a uniform sampling procedure that we might be
able to use for this purpose.
To evaluate the soundness of the method described in this publication, we could
compare the predictive ability obtained with 'snp77' from uniform sampling with
the mean (or median) of the predictive ability for 'snp77' from repeated random
sampling (20 replications, see Figure
\@ref(fig:Resampled-Predictive-Ability-Plots)).

## Algorithm validation
To validate the algorithms from @Fernando2014 and its extension in
\@ref(introduction), we could perhaps have a look at an independent data set,
such as the one described by @Guo2016, who used data from the [Yan
lab](http://www.maizego.org/Resources.html).