### 2017-03-14 reports/index.Rmd
Remove chapter overview from report index

---
title: Single-step Hybrid Predition
author:
  - name: Matthias Westhues
    email: matthias.westhues@uni-hohenheim.de
    affiliation: University of Hohenheim
  - name: Claas Heuer
    email: cheuer@tierzucht.uni-kiel.de
    affiliation: University of Kiel
  - name: Georg Thaller
    email: thaller@tierzucht.uni-kiel.de
    affiliation: University of Kiel
  - name: Rohan Fernando
    email: rohan@iastate.edu
    affiliation: Iowa State University
  - name: Albrecht E. Melchinger
    email: melchinger@uni-hohenheim.de
    affiliation: University of Hohenheim
    footnote: Corresponding Author
address:
  - code: University of Hohenheim
    address: Institute of Plant Breeding, Seed Science and Population Genetics, Fruwirthstr. 21, 70599 Stuttgart, Germany
  - code: University of Kiel
    address: Institut of Animal Breeding and Husbandry, Hermann-Rodewald-Straße 6, 24118 Kiel, Germany
  - code: Iowa State University
    address: Department of Animal Science, 239 Kildee Hall, 50011 Ames, Iowa, U.S.A.
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
bibliography: ["library.bib", "cheuer.bib"]
abstract: |
  This is the abstract.

  It consists of two paragraphs.
---