### 2017-03-14 reports/methods.Rmd
Remove analyses from methods

# Materials and Methods

## Data
<!--
./reports/methods_data.R
-->

### Agronomic dat
The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
mating designs between 142 Dent and 103 Flint parent lines [@Westhues2017]. 
Evaluations of six agronomically important traits were recorded on all hybrids
in field experiments at three or more agro-ecologically diverse locations 
across Germany.

The maize inbred line data set originally comprised 513 maize lines representing 
the global maize diversity and was reduced to the set of tropical and
subtropical lines ($n = 211$) given that it is the largest of the four
pre-classified subgroups [@Yang2014].
All inbred lines were evaluated in five Chinese environments ranging from
$18^{\circ}$ to $30^{\circ}$N and from $102^{\circ}$ to $110^{\circ}$E
[@Yang20114].
Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
we will evaluate six in our prediction models.

#### Predictor data
All parental inbred lines of the maize hybrids were genotyped using the Illumina 
SNP BeadChip MaizeSNP50 [@Ganal2011].

MaizeSNP50 BeadChip-data were available for the entire set of lines from the
diversity panel.
Additionally, RNA sequencing was performed for 368 lines on young seedlings 15 
days after sowing.
By exploting identify by descent (IBD) between 49,728 SNPs on the BeadChip 
overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
for these lines [@Fu2013,Li2013].
For the remaining 145 maize lines without RNA-seq data, high density markers
were inferred via projection of IBD regions onto the BeadChip, using this core 
set of markers [@Yang2014].
Gene expression data were normalized using a normal quantile transformation to
satisfy modelling assumptions [@Fu2013].

After filtering for minor allele frequency ($ \geq 5\%$), heterozygosity rate
($\geq 5\%$) and call frequency ($\geq 95\%$), missing genotypes were imputed 
using the Beagle software [@Browning2009].
The final number of polymorphic marker loci was 7,013 for parental Dent and
6,212 for parental Flint lines, respectively.

### Pedigree Data (Hybrid)

## Structure

### PCA

### Population Structure Analysis

## Core Sampling

## Prediction

### Kernels

### Imputation

### Models

### Leave-one-out cross-validation (LOOCV)