### 2017-03-14 reports/methods.Rmd
Describe UHOH predictor data

# Materials and Methods

## Data
<!--
./reports/methods_data.R
-->

### Hybrid data
The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
mating designs between 142 Dent and 103 Flint parent lines [@Westhues2017]. 
Best linear unbiased estimates (BLUEs) were computed across all factorials and 
three or more agro-ecologically diverse locations across Germany for six 
agronomically important traits.
All 245 parent lines have pedigree records back to the generation of their 
grandparents or deeper [@Westhues2017] and were genotyped using the Illumina 
SNP BeadChip MaizeSNP50 [@Ganal2011].
After filtering for minor allele frequency ($ \geq 5\%$), heterozygosity rate
($\geq 5\%$) and call frequency ($\geq 95\%$), missing genotypes were imputed 
using the Beagle software [@Browning2009].
The final number of polymorphic marker loci was 7,013 for parental Dent and
6,212 for parental Flint lines, respectively.
Gene expression data for seedlings seven days after sowing were obtained for a 
subset of 60 parental Dent and 43 parental Flint lines through two-color
hybridizations using a custom 2K microarray (GPL22267) [@Westhues2017].
After applying established normalization procedures [@Smyth2003,Ritchie2007], 
gene expression BLUEs were computed for 1,323 transcripts in reference to 
established protocols [@Symth2003,Ritchie2007,Frisch2010] using the R-package 
*limma* [@Ritchie2015].

#### Diversity panel
The maize inbred line data set originally comprised 513 maize lines representing 
the global maize diversity and was reduced to the set of tropical and
subtropical lines ($n = 211$) given that it is the largest of the four
pre-classified subgroups [@Yang2014].
All inbred lines were evaluated in five Chinese environments ranging from
$18^{\circ}$ to $30^{\circ}$N and from $102^{\circ}$ to $110^{\circ}$E
[@Yang20114].
Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
we will evaluate six in our prediction models.
MaizeSNP50 BeadChip-data were available for the entire set of lines from the
diversity panel.
Additionally, RNA sequencing was performed for 368 lines on young seedlings 15 
days after sowing.
By exploting identify by descent (IBD) between 49,728 SNPs on the BeadChip 
overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
for these lines [@Fu2013,Li2013].
For the remaining 145 maize lines without RNA-seq data, high density markers
were inferred via projection of IBD regions onto the BeadChip, using this core 
set of markers [@Yang2014].
The same SNP quality checks as for the parental inbred lines of the hybrids
were applied to the 211 inbred lines from the tropical/subtropical subset,
yielding 37,760 SNPs.
Gene expression data were normalized using a normal quantile transformation to
satisfy modelling assumptions [@Fu2013].
A subset of 28,850 annotated genes was kept for further analyses [@Li2013].

### Pedigree Data (Hybrid)

## Structure

### PCA

### Population Structure Analysis

## Core Sampling

## Prediction

### Kernels

### Imputation

### Models

### Leave-one-out cross-validation (LOOCV)