### 2017-03-14 reports/methods.Rmd
Methods: Core sampling, PCA, Admixture

# Materials and Methods

## Data
<!--
./reports/methods_data.R
-->

### Hybrid data
The maize hybrid data set comprises 1,521 hybrids produced in 16 factorial
mating designs between 142 Dent and 103 Flint parent lines [@Westhues2017]. 
Best linear unbiased estimates (BLUEs) were computed across all factorials and 
three or more agro-ecologically diverse locations across Germany for six 
agronomically important traits.
All 245 parent lines have pedigree records back to the generation of their 
grandparents or deeper [@Westhues2017] and were genotyped using the Illumina 
SNP BeadChip MaizeSNP50 [@Ganal2011].
After filtering for minor allele frequency ($ \geq 5\%$), heterozygosity rate
($\geq 5\%$) and call frequency ($\geq 95\%$), missing genotypes were imputed 
using the Beagle software [@Browning2009].
The final number of polymorphic marker loci was 7,013 for parental Dent and
6,212 for parental Flint lines, respectively.
Gene expression data for seedlings seven days after sowing were obtained for a 
subset of 60 parental Dent and 43 parental Flint lines through two-color
hybridizations using a custom 2K microarray (GPL22267) [@Westhues2017].
After applying established normalization procedures [@Smyth2003,Ritchie2007], 
gene expression BLUEs were computed for 1,323 transcripts in reference to 
established protocols [@Symth2003,Ritchie2007,Frisch2010] using the R-package 
*limma* [@Ritchie2015].

#### Diversity panel
The maize inbred line data set originally comprised 513 maize lines representing 
the global maize diversity and was reduced to the set of tropical and
subtropical lines ($n = 211$) given that it is the largest of the four
pre-classified subgroups [@Yang2014].
All inbred lines were evaluated in five Chinese environments ranging from
$18^{\circ}$ to $30^{\circ}$N and from $102^{\circ}$ to $110^{\circ}$E
[@Yang20114].
Best linear unbiased predictors (BLUPs) were calculated for 17 traits of which 
we will evaluate six in our prediction models.
MaizeSNP50 BeadChip-data were available for the entire set of lines from the
diversity panel.
Additionally, RNA sequencing was performed for 368 lines on young seedlings 15 
days after sowing.
By exploting identify by descent (IBD) between 49,728 SNPs on the BeadChip 
overlapping with the RNA-seq data, 556,809 high quality SNPs could be inferred 
for these lines [@Fu2013,Li2013].
For the remaining 145 maize lines without RNA-seq data, high density markers
were inferred via projection of IBD regions onto the BeadChip, using this core 
set of markers [@Yang2014].
The same SNP quality checks as for the parental inbred lines of the hybrids
were applied to the 211 inbred lines from the tropical/subtropical subset,
yielding 37,760 SNPs.
Gene expression data were normalized using a normal quantile transformation to
satisfy modelling assumptions [@Fu2013].
A subset of 28,850 annotated genes was available for 149 out of the 211
genotypes and was kept for further analyses [@Li2013].

## Population structure
Principal component analyses (PCA) were run on scaled and centered predictor 
matrices using the *pca()* function from the R-package *LEA* [@Frichot2015].
Estimates of ancestry coefficients were estimated based on parse nonnegative
matrix factorization algorithms [@Frichot2014] using the *snmf()* function 
in *LEA*.
Algorithms were run for $K=2, \dots 5$ putative ancestral gene pools with 25
repetitions for each value of $K$. 
For each $K$, only the repetition with the lowest cross-entropy was kept for 
further analyses.

## Core Sampling
In order to evaluate the influence of the genetic constitution of the set of
genotypes that has only data on one out of two predictors, we generated core 
samples using the R-package *corehunter* [@Thachuk2009].
These core samples were based on the 149 maize inbred lines from the diversity
panel which had gene expression as well as genomic information.
The size of the core set was varied in increments of ten percentage points and 
ranged from 10% to 90% of all 149 inbred lines.
For assembling the various core sets, we first computed the Modifed Rogers (MR)
distance between each pair of genotypes and then aimed at maximizing the
average distance between each member of the selected core.
MR was computed using the *rogers.dist()*-function from the R-pacakge *poppr*
[@Kamvar2015].

## Prediction

### Kernels

### Imputation

### Models

### Leave-one-out cross-validation (LOOCV)