### 2017-03-15 paper/genetics.tex
Init LaTeX paper

\documentclass[12pt,titlepage]{article}

%% THE USEPACKAGES NECESSARY FOR THIS EXAMPLE
%% NOTE THAT genetics_manu_style MUST BE CALLED AFTER mychicago
\usepackage{graphicx}
\usepackage{endfloat}
\usepackage{amsfonts}
\usepackage{mychicago}
\usepackage{subfigure}
\usepackage{genetics_manu_style}
\usepackage{authblk} % improved citations
\usepackage{setspace} %double spacing
\usepackage{lineno} % line numbers
\usepackage{booktabs} % book-quality tables
\usepackage{multicol} % list corresponding authors in two separate columns

%% Probably unnecessary packages.
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript

\usepackage{xr}
\externaldocument{supp_info}

%% THE MANUSCRIPT TITLE
\title{Omics-based Hybrid Prediction in Maize}

% Allow for multiple authors to share the same institution.
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

\author{
  Matthias Westhues\thanks{equally contributing}\thanks{Institute of Plant Breeding, Seed Science and Population Genetics, University of Hohenheim, D-70599 Stuttgart, Germany},
  Tobias A. Schrag\samethanks[1]\samethanks[2],
  Claas Heuer\thanks{Institute of Animal Breeding and Husbandry, Christian-Albrechts-University Kiel, D-24098 Kiel, Germany},
  Georg Thaller\samethanks[3],
  H. Friedrich Utz\samethanks[2],
  Wolfgang Schipprack\samethanks[2],
  Alexander Thiemann\thanks{Biocenter Klein Flottbeck, Developmental Biology and Biotechnology, University of Hamburg, D-22609 Hamburg, Germany},
  Felix Seifert\samethanks,
  Anita Ehret\samethanks[3],
  Armin Schlereth\thanks{Max-Planck Institute of Molecular Plant Physiology,
  D-14476 Potsdam, Germany},
  Mark Stitt\samethanks,
  Zoran Nikoloski\samethanks,
  Lothar Willmitzer\samethanks,
  Chris C. Sch{\"o}n\thanks{Plant Breeding, Technische Universit{\"a}t
  M{\"u}nchen, D-85354 Freising, Germany},
  Stefan Scholten\samethanks[4],
  Albrecht E. Melchinger\samethanks[2]
}

\renewcommand{\CorrespondingAddressA}{%
    University of Hohenheim \\
    Institute of Plant Breeding, Seed Science and Population Genetics \\ 
    Fruwirthstr. 21 \\ 
    Stuttgart 70599, GERMANY \\
    Tel.: (+49) 0711 459-22334 \\
    Fax: (+49) 0711 459-22343 \\ 
    \texttt{melchinger@uni-hohenheim.de} \vfill}

\renewcommand{\CorrespondingAddressB}{%
    University of Hamburg \\
    Biocenter Klein Flottbeck, Developmental Biology and Biotechnology \\ 
    Ohnhorststr. 18 \\ 
    Hamburg 22609, GERMANY \\ 
    Tel.: (+49) 40 42816-286 \\
    Fax: (+49) 40 42816-252 \\ 
    \texttt{stefan.scholten@uni-hamburg.de} \vfill}

\renewcommand{\RunningHead}{Omics-based Hybrid Prediction}
\renewcommand{\CorrespondingAuthorA}{Prof. Dr. A. E. Melchinger}
\renewcommand{\CorrespondingAuthorB}{PD Dr. Stefan Scholten}
\renewcommand{\KeyWords}{omics, hybrid breeding, prediction of complex traits, maize, endophenotype}

%% SOME COMMANDS FOR THE CONTENT OF THIS FILE.  NOT NECESSARY FOR
%% GENETIC_MANU_STYLE
\newcommand{\bp}{\mathbf{p}}
\newcommand{\LLL}{\mathcal{L}}

%% BEGIN DOC
\begin{document}

\maketitle
\doublespacing
\linenumbers

\begin{abstract}
Accurate prediction of traits with complex genetic architecture is crucial for 
selecting superior candidates in animal and plant breeding and for guiding 
decisions in personalized medicine.
Whole-genome prediction (WGP) has revolutionized these areas but has inherent 
limitations in incorporating intricate epistatic interactions.
Downstream 'omics' data are expected to integrate interactions within and 
between preceding biological strata and provide the opportunity to improve trait 
prediction.
Yet, predicting traits from parents to progeny has not been addressed by a 
combination of 'omics' data. 
Here, we evaluate several 'omics' predictors --- genomic, transcriptomic and 
metabolic data --- measured on parent lines at early developmental stages, and 
demonstrate that the integration of transcriptomic with genomic data leads to 
higher success rates in the correct prediction of untested hybrid progeny in 
maize.
Despite the high predictive ability of genomic data, transcriptomic data alone 
outperformed them and other predictors for the most complex heterotic trait, 
dry matter yield.
An eQTL analysis revealed that transcriptomic data integrate genomic information
from both, adjacent and distant sites relative to the expressed genes.
Together, these findings suggest that downstream predictors capture
physiological epistasis that is transmitted from parents to their hybrid
offspring.
We conclude that the use of downstream 'omics' data in prediction can exploit 
important information beyond structural genomics for leveraging the efficiency 
of hybrid breeding.
\end{abstract}

% Please add here a significance statement to explain the relevance of your work
\section{Article Summary}
The efficiency of modern hybrid breeding programs could be greatly improved
by pre-selecting the very best out of millions of theoretically possible
candidates using molecular 'omics'-features of their parents.
% Data for the following statement from 
% '/reduced_pred/100_Combined_Analyses/integrate.R'
Here, we show that transcriptomic data alone increased the predictive ability
for yield by 14.7\% compared to the gold standard, genomic data.
This improvement translates into a substantial increase in the probability of
success for predicting the best hybrids.
A combination of transcriptomic with genomic data yielded optimal and stable
predictive abilities across a wide range of traits.

% Introduction
\section{Introduction} 
Hybrid breeding, which entails crossing of lines from two genetically distant
germplasm collections --- called heterotic groups \cite{Melchinger1998} --- 
has emerged as a prime strategy to meet demands for a sustainable
intensification of agricultural production \cite{Duvick2005}.
However, unlocking the full potential of hybrid breeding requires accurate
prediction methods to efficiently identify the superior candidates out of the
millions of possible hybrids that could potentially be produced in each
cycle of an ordinary-sized breeding program.
With the advent of the doubled haploid (DH) technology \cite{Wedzony2009}
this prediction problem has become even more challenging because, based on 
breeder's experience, the vast majority ($\approx 90$\%) of competing lines in 
each heterotic group are "new" lines without any phenotypic records on hybrid
progeny from previous breeding cycles.
Consequently, among all hybrid combinations possible between lines from two
heterotic groups, about 81\% are T0 hybrids, 18\% are T1 hybrids and 1\% are T2
hybrids having zero, one or two parents, respectively, that have been previously
tested in other hybrid combinations.
Preselection of a few hundred of the most favorable hybrids with high success
rate could significantly reduce the labor-intensive and time-consuming
field-testing \cite{Kadam2016,Xu2016}.
This could greatly impact the efficiency of hybrid breeding and boost the
annual selection gain \cite{Longin2015}.

Whereas yield and other heterotic traits of hybrids are generally poorly
predicted by the performance of their parent lines \cite{Melchinger1998}, WGP
has emerged as a major tool for tackling this challenge
\cite{Massman2013,Technow2014}.
Nevertheless, there is evidence that, even with complete sequence information,
genomic prediction may not capture complex interactions between biological 
strata or downstream regulation that act through the entire cascade from 
genotype to phenotype \cite{Dalchau2011,Zhu2012,Rudd2015,Ritchie2015b}.
Most studies have evaluated predictive ability by looking at only one kind of 
endophenotype (intermediaries between genotype and phenotype 
\cite{Gottesman2003,Mackay2009}) such as the transcriptome 
\cite{Swanson-Wagner2006,Frisch2010,Xu2016} or the metabolome 
\cite{Meyer2007,Xu2016,Dan2016}.
The integration of different endophenotypic and genomic data is expected to
reflect more closely the phenotypic variability across individuals 
\cite{Mackay2009,Patti2012,Civelek2014}. 
Two recent studies that integrated multiple biological strata in predicting 
breast cancer risk \cite{Vazquez2016} and performance of maize inbred lines
\cite{Guo2016} demonstrated the benefit of this strategy.
However, unlike forecasting clinical or agronomic traits from endophenotypes of 
the same individual, hybrid breeding requires the prediction of the genotypic 
values (GV) of hybrid progeny based on parental information.
To achieve this objective, we used the BLUP approach --- originally developed in
animal breeding \cite{Henderson1984} --- for the more complex setting of hybrids 
between parents from two heterotic groups 
\cite{Bernardo1994,Bernardo1995,Bernardo1996,Massman2013}.
Here, we measured endophenotypes of parent lines to forecast the GV of
T0, T1 and T2 hybrid progeny by using prediction equations trained with "omics" 
information from related parent lines and phenotypic information on their
hybrid offspring.

\section{Materials and Methods}
\subsection{Genetic material and phenotyping}
The entire genetic material consisted of a set of 1,536 hybrids, denoted as 
$H_{Tot}$, produced in 16 factorial mating designs between 142 Dent and 103 
Flint lines from the maize breeding program at the University of Hohenheim, on 
which agronomic data for silage maize production, as well as pedigree and 
genomic data were available.
A subset of this material, albeit from different trials, has been used for
genomic prediction of traits related to grain maize production
\cite{Technow2014}.
For hybrid prediction, we used a core set $H \subset H_{\text{Tot}}$ of 617 
hybrids, produced in six factorials with hybrid sets $H_{\text{FAC}^{(i)}} (i =
1, 2, \dots, 6; H = \bigcup_{i=1}^{6} H_{\text{FAC}^{(i)}})$ from crosses 
between 57 Dent and 41 Flint inbred lines, denoted as $D = \{1, 2, \dots, 57 \}$ 
and $F = \{1, 2, \dots, 41 \}$ (File S1).
All hybrids were evaluated in field experiments at three or more 
agro-ecologically diverse locations across Germany.
In the trials of each factorial, which included at least five common check 
genotypes, the entries were randomized in $\alpha$ lattice designs and planted 
in 2-row plots.
Dry matter yield (DMY, $t/ha$) and dry matter content (DMC, \%) of whole-plant
aboveground biomass were determined by established procedures 
\cite{Riedelsheimer2012b}. 
For quality traits, contents of fiber (ADF, \%), fat (FAT, \textperthousand), 
protein (PRO, \textperthousand), starch (STA, \%), and sugars (SUG,
\textperthousand) in dry matter were measured in the harvested plant material 
using calibrated near-infrared spectroscopy (NIRS; \citeNP{Grieder2011}, 
File S1).

\subsection{Pedigree-based relationship coefficients}
Pedigree records were gathered for all parent lines at least back to the 
generation of their grandparents.
Coancestry coefficients were calculated using SAS (version 9.4, SAS Institute) 
for all possible pairs of lines in each heterotic group according to established 
rules \cite{Falconer1996} under the following assumptions \cite{Cox1986}:
(i) all lines in a pedigree are genetically homogeneous and homozygous, (ii) 
pairs of genotypes with no known common parentage are unrelated, and (iii) a 
line derived from a cross or backcross obtained a proportional fraction of the 
genome from each parent, as expected under Mendelian inheritance in the absence 
of selection.

\subsection{Genotyping}
Genotyping of all inbred lines was performed with the Illumina SNP chip
MaizeSNP50 \cite{Ganal2011}.
After performing a commonly used quality check \cite{Technow2014} and 
imputation of missing data \cite{Browning2009}, a total of 21,565 polymorphic 
SNPs was available and used for all further analyses.

\subsection{Metabolite profiling}
Seedlings of all parental inbred lines were grown under controlled conditions 
inside climate chambers to quantify their root metabolite profiles as detailed
by \citeN{DeAbreueLima2017}.
The experiment was laid out as a randomized incomplete block design with
replicated germination boxes.
For leaf metabolic profiles (known metabolites and unannotated chromatographic
peaks), a field experiment was carried out in an $\alpha$ lattice design with 
two replications at one location in southern Germany in the spring of 2012.
Excision of leaves at the third leaf stage was performed according to an 
established protocol \cite{Riedelsheimer2012b} and finalized within 45 minutes 
for the entire experiment. 
For both profiling procedures, all material was transferred directly into 
containers with dry ice and then into liquid nitrogen to quench metabolic 
activity.

\subsection{Transcriptome profiling}
For transcriptome profiling, five seeds per parent line were taken from the 
same seed lot as used for metabolite profiling and laid out inside a climate 
chamber in a randomized complete block design with five replications.
Seedlings were sampled seven days after sowing, snap-frozen in liquid nitrogen,
and stored at -80$^{\circ}$C until use.
Prior to mRNA extraction, all replicates of a genotype were pooled and
homogenized.
A custom 2K-microarray (GPL22267) was assembled from a subset of the 47K maize 
oligonucleotide array (GPL6438).
Two-color hybridizations were carried out separately for each of the six 
factorials using interwoven loop designs \cite{Kerr2001}.
The average number of shared genotypes between factorials was 4.5 and ranged 
from 2 to 10.

\subsection{Statistical analysis of agronomic traits}
Agronomic data were analyzed in two stages, following \citeN{Technow2014}. 
In the first stage, and separately for each environment, adjusted-entry means 
were computed for every hybrid using REML-based linear mixed-model analyses.
In the second stage, best linear unbiased estimates (BLUEs) were computed for 
all hybrids in $H_{Tot}$.
The BLUEs of hybrids in the core set $H$ served as response variables in our 
hybrid prediction models and cross-validation routines. 
For all predictions we used computationally efficient best linear unbiased 
predictor (BLUP) models, which have the same properties as those of a selection 
index because we previously accounted for fixed effects 
(\citeNP{Mrode2014}, pp. 34, 311, 312).
For general and specific combining abilities (GCA and SCA) of parent lines,
we used ASReml \cite{asreml} to compute best linear unbiased predictors (BLUPs), 
variance components ($\sigma^2_{\text{GCA}^{D}}$, $\sigma^2_{\text{GCA}^{F}}$, 
$\sigma^2_{\text{SCA}}$) and entry-mean heritabilities ($H^2$) of all hybrids
in $H_{Tot}$, treating all effects in the model as random.
The covariance matrices of the GCA and SCA effects were defined by multiplying
the variance components with their respective genomic relationship matrices 
(File S1).

\subsection{Statistical analysis of endophenotypes}
Raw data were normalized using established procedures for metabolites
\cite{VandenBerg2006} and transcripts \cite{Smyth2003,Ritchie2007}.
From these data, we obtained BLUEs for metabolite levels and transcript 
abundance of each line using REML-based mixed-model analyses.
The statistical models for the analysis of metabolite profiles accounted for 
various experimental effects as detailed by \citeN{DeAbreueLima2017}.
After applying quality checks and computing BLUEs, 92 leaf metabolic analytes 
and 283 root metabolic analytes remained for further analyses.
BLUEs for transcriptomic data were computed using the R-package \emph{limma} 
\cite{Ritchie2015a} in reference to established protocols 
\cite{Smyth2003,Ritchie2007,Frisch2010}.
The design matrix for the linear model was based on the dye-labeling of a 
reference genotype and included a fixed effect for each factorial design 
in order to account for the use of biological replicates of some genotypes in 
different arrays. 
All gene expression values were subsequently computed, based on the log-ratio 
relative to this common genotype \cite{Smyth2004}.
In total, 1,323 gene expression profiles were available.
Repeatabilities ($w^2$) were estimated for each endophenotype at the inbred line 
level using the same models as for the computation of BLUEs, but treating the 
genotype effect as random.
This analysis was performed jointly for the Dent and Flint lines allowing for
different means and heterogeneous genotypic variances of the heterotic groups,
but assuming a common error variance.
Variance components were estimated by Gibbs sampling using the \emph{R} package
\emph{MCMCglmm} \cite{Hadfield2010}.

\subsection{Prediction models and model evaluation}
Predictions of hybrid performance were compared on the basis of the core set
of hybrids $H$ and the corresponding sets of parent lines $D$ and $F$ on which
data for all five predictors (P, pedigree; G, genomic; T, transcriptomic; L,
leaf metabolic; R, root metabolic data) were available with the exception of
missing data in a few lines for R.
The matrices $\mathbf{W}_{D}$ and $\mathbf{W}_{F}$ are matrices of standardized 
feature measurements for the various predictors (G, T, L, R).
The matrix $\mathbf{W}$ has dimension 'number of parent lines in the 
corresponding heterotic group' times 'number of features'.
The columns in $\mathbf{W}_{D}$ and $\mathbf{W}_{F}$ are centered and
standardized to unit variance, respectively.

The kernels pertaining to each predictor and lines from each heterotic group
--- corresponding to genomic relationship matrices in the case of SNPs --- can 
then be defined as
\begin{equation} \label{eq:GenomicRelationship}
  \mathbf{G}_{D} = \frac{1}{W} \mathbf{W}_{D} \mathbf{W}_{D}^{\top}, \
  \mathbf{G}_{F} = \frac{1}{W} \mathbf{W}_{F} \mathbf{W}_{F}^{\top},
\end{equation}

where $W$ denotes the number of features \cite{VanRaden2008}.
In the case of pedigree data (P), coancestry coefficients were used directly
for $\mathbf{G}_{D}$ and $\mathbf{G}_{F}$, respectively.

The universal model for GCA and SCA effects was:

\begin{equation} \label{eq:HypredDominance}
  \mathbf{y} = \mu + 
  \sum_{p=1}^{P} \mathbf{Z}_{D} \mathbf{g}_{Dp} +
  \sum_{p=1}^{P} \mathbf{Z}_{F} \mathbf{g}_{Fp} +
  \sum_{p=1}^{P} \mathbf{Z}_{S} \mathbf{s}_{p} +
  \mathbf{\epsilon},
\end{equation}

where $\mathbf{y}$ is the vector of observed hybrid performance (BLUEs), $\mu$ 
is the fixed model intercept, $\mathbf{Z}_{D}$ is the
corresponding design matrix associating the random GCA effects of the lines in
$D$ ($\mathbf{g}_{Dp}$) with $\mathbf{y}$, $\mathbf{Z}_{F}$ is the 
corresponding design matrix associating the random GCA effects of the lines in 
$F$ ($\mathbf{g}_{Fp}$) with $\mathbf{y}$ and $\mathbf{Z}_{S}$ is a design 
matrix associating the SCA effects ($\mathbf{s}_{p}$), pertaining to hybrid 
combinations for the $p$-th predictor data type with the corresponding hybrid 
measurements in $\mathbf{y}$.
The random effects ($\mathbf{g}_{Dp}$ and $\mathbf{g}_{Fp}$) have expectation 
zero and variance equal to $\mathbf{G}_{Dp} \sigma^{2}_{\text{GCA}^{Dp}}$ and 
$\mathbf{G}_{Fp} \sigma^{2}_{\text{GCA}^{Fp}}$for the GCA effects of the Dent
and Flint lines, respectively,
$\mathbf{S}_{p} \sigma^{2}_{\text{SCA}^{p}}$ for the SCA effects and
$\mathbf{I} \sigma^2_{\epsilon}$ for the residual error.
For each pair of hybrids $i \times k$ and $j \times l$, the corresponding
elements in $\mathbf{S}_{p}$ were obtained as the product of the respective
elements $f_{ij}$ in $\mathbf{G}_{Dp}$ and $f_{kl}$ in $\mathbf{G}_{Fp}$,
respectively \cite{Schnell1965,Bernardo1996,Massman2013} (File S1).
In the absence of epistasis, this model is equivalent to a feature model
accounting for dually defined additive effects in each heterotic group and
dominance effects between them.
Extensions of the single-predictor models were made by adding GCA and SCA
effects for any additional predictor assuming stochastic independence of
effects.
In order to obtain unbiased estimates of the predictive ability and to compare
different models and predictor combinations, following \citeN{Technow2014}, we 
devised a cross-validation (CV) scheme, stratified by the parent lines and using 
1,000 runs (CV1000, File S1). All prediction models were implemented using the 
\emph{R} package \emph{BGLR} \cite{Perez2014}.

\subsection{Comparison of predictive abilities}
Predictive abilities were obtained by calculating Pearson correlations between
predicted ($\hat{y}$) and observed phenotypes ($y$), separately for three
test set partitions (T0, T1 and T2 hybrids).
For each CV run, the training and validation sets were stored to ensure the
validity of comparisons between any predictor and combinations thereof.
For any two predictors, say $A$ and $B$, we have the corresponding vectors 
with predictive abilities $p_{A}$ and $p_{B}$ of length 'number of cross
validation runs'.
For comparing predictive abilities obtained for different predictors and models,
we first ensured that the difference between any two scenarios (\textit{i.e.}, 
$p_{A} - p_{B}$) is normally distributed \cite{Dietterich1998}.
Then, we applied multiple contrast tests on the mean of the distribution of 
predictive abilities, obtained from the 1,000 CV runs, using the \emph{R} 
package \emph{multcomp} \cite{Hothorn2008} and Tukey contrasts.
This procedure is similar to the pairwise \textit{t}-test \cite{Guo2016} but 
accounts for multiple comparisons.

\subsection{Evaluation of a pre-selection bias in transcriptomic data}
A custom 2K-microarray (GPL22267) was assembled from a subset of the 47K maize 
oligonucleotide array (GPL6438), based on association of genes with hybrid
performance or mid-parent heterosis for grain yield and grain dry matter
content of maize. 
These two traits were evaluated in separate grain-yield trials with hybrids from 
factorial $H_{\text{FAC}^{(1)}}$ (\citeNP{Frisch2010,Thiemann2010}, File S1). 
To ensure that no pre-selection bias was introduced in hybrid prediction using 
these transcriptomic data, we compared predictive abilities for $H$ with 
predictive abilities when excluding $H_{\text{FAC}^{(1)}}$.

\subsection{Association mapping}
For each of the seven agronomic traits, we performed a genome-wide association
study (GWAS) with GCA effects of all 142 Dent and 103 Flint parent lines as
response variables using the EMMAX-method \cite{Kang2010} as implemented in 
\emph{cpgen} \cite{Heuer2015}.
To avoid using the marker data twice, GCA effects were calculated using only
pedigree information.
Furthermore, an eQTL analysis was carried out to examine statistically 
significant associations between genomic and transcriptomic data for the parent
lines ($D$ and $F$) of the core set $H$ plus five additional lines.
This was accomplished in the same way as in the GWAS for agronomic traits, but
here the BLUPs of the transcriptomic data of each mRNA were used as the 
response variables.
Associations in each GWAS were declared statistically significant at 
$\alpha = 0.05$ after Bonferroni correction.

\subsection{Probability of success}
Calculations of the conditional probability ($P[r, \beta]$) of a hybrid, 
selected at random from the upper $\beta$ percent fraction of the distribution 
of predicted values having a phenotypic value contained in the upper $\beta$ 
percent of the distribution of observed phenotypic values, assumed a bivariate
normal distribution

\begin{align}
  \begin{bmatrix}
    \hat{y} \\
    y
   \end{bmatrix}
  & \sim
  N \Bigg(
  \begin{bmatrix}
    0 \\
    0
  \end{bmatrix},
  \begin{bmatrix}
    1 & r \\
    r & 1
  \end{bmatrix}
  \Bigg).
\end{align}

and were performed within the \emph{R} statistical environment to solve the 
required integrals \cite{Robson1967}.

\subsection{Principal component analysis}
For principal component analyses of predictor data, all variables were scaled 
and centered.
Clusters represent two component mixtures of bivariate t-distributions, which
were estimated using Maximum Likelihood.
Ellipses were drawn based on the 0.95 quantiles of the respective bivariate
t-distributions.
Unless stated otherwise, all statistical analyses were carried out inside the 
\emph{R} environment for statistical computing \cite{Team2016}.

\subsection{Data availability}
Adjusted means of hybrid performances are provided in File S2, genomic marker
data are provided in File S3, genomic marker positions are provided in File
S4, transcriptomic BLUEs are provided in File S5, leaf metabolic BLUEs are
provided in File S6 and root metabolic BLUEs are provided in File S7.
Genotypes can be assigned to either the Dent or the Flint heterotic group using
a classification table provided in File S8.
Code used to analyze the data is available upon request.

\section*{Results}
\Genetics2level{Agronomic data}
Mean values of the 1,536 hybrids for the seven evaluated agronomic traits, with 
relevance to animal feed and biogas production, were of the same magnitude as 
reported by \citeN{Riedelsheimer2012b} and \citeN{Grieder2011}.
For all traits, $\sigma^{2}_{\text{GCA}^{D}}$ and
$\sigma^{2}_{\text{GCA}^{F}}$, describing the main effects of the parents from
each heterotic group, together explained more than 93\% of the genotypic
variance among hybrids (Table \ref{table:VarianceComponents}).
Heritabilities were moderate to high for all agronomic traits, indicating a 
high precision of field experiments and data collection 
(Table \ref{table:VarianceComponents}).

% -- TABLE 1
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[tbhp]
\centering
\caption{
  Summary of agronomic traits.
  Traits are characterized by overall mean ($\mu$), variance components of GCA
  effects for Dent ($\sigma^2_{\text{GCA}^D}$) and Flint lines
  ($\sigma^{2}_{\text{GCA}^{F}}$) and SCA effects ($\sigma^{2}_{SCA}$)
  (followed by s.e.m.) as well as entry mean heritabilities ($H^2$).
}
\label{table:VarianceComponents}
\begin{tabular}{@{}llllll@{}}
\toprule
Trait                   & $\mu$ & $\sigma^{2}_{\text{GCA}^{D}}$ & $\sigma^{2}_{\text{GCA}^{F}}$ & $\sigma^{2}_{\text{SCA}}$ & $H^{2}$         \\ \midrule
  DMY (t/ha)                    & 19.00 & $1.51 \pm 0.22$               & $1.00 \pm 0.18$               & $0.17 \pm 0.03$           & $0.82$ \\
  DMC (\%)                    & 34.13 & $4.17 \pm 0.59$               & $5.03 \pm 0.79$               & $0.49 \pm 0.07$           & $0.91$ \\
  ADF (\%)                    & 20.93 & $0.27 \pm 0.06$               & $0.40 \pm 0.09$               & $0.02 \pm 0.02$           & $0.43$ \\
  FAT (\textperthousand)                    & 30.02 & $1.01 \pm 0.19$               & $2.10 \pm 0.37$               & $0.17 \pm 0.05$           & $0.73$ \\
  PRO (\textperthousand)                    & 69.65 & $3.11 \pm 0.53$               & $2.77 \pm 0.51$               & $0.29 \pm 0.10$           & $0.70$ \\
  STA (\%)                    & 35.56 & $2.95 \pm 0.51$               & $3.33 \pm 0.63$               & $0.24 \pm 0.10$           & $0.69$ \\
  SUG (\textperthousand)                    & 38.12 & $31.33 \pm 5.15$              & $31.90 \pm 5.73$              & $4.12 \pm 1.02$           & $0.77$ \\ \bottomrule
\end{tabular}
\end{table}

\Genetics2level{Predictor data}
% Repeatabilities
Repeatabilities ($w^{2}$) for endophenotypes varied considerably in both groups
of parents (Fig. S\ref{fig:Repeatability}\textit{A}) with average values
ranging from 0.31 to 0.41, except for transcriptomic data in Flint material 
where the average repeatability was only 0.18.
Nevertheless, in the latter case, 291 out of 1,323 transcripts still
exceeded a threshold of $w^{2} = 0.4$.

Dent and Flint lines were clearly separated in principal component analyses of 
genomic and transcriptomic data (Fig. \ref{fig:PredictorProperties}\textit{A}).
However, they overlapped for leaf metabolic and, to an even greater extent, for
root metabolic data.
Off-diagonal elements of the kernels $\mathbf{G}_{D}$ and $\mathbf{G}_{F}$, 
respectively, showed moderate correlations between genomic and transcriptomic 
data ($\rho_{D} \approx 0.56, \rho_{F} \approx 0.44$, 
Fig. S\ref{fig:Rel_Assoc}). 
Correlations between the off-diagonal elements of the $\mathbf{G}$-matrices were 
highest for the comparison between genomic and pedigree data 
($\rho_{D} \approx 0.72$, $\rho_{F} \approx 0.63$).
Intriguingly, the associations between the $\mathbf{G}$-matrices for the root
and leaf metabolic data were very low ($\rho_{D} \approx 0.12$,
$\rho_{F} \approx 0.06$).

% Figure 1
\begin{figure}[H]
\centering
\includegraphics[width=0.90\linewidth]{./figures/predictor_properties_vertical.pdf}
\caption{
  Properties of predictor data for Dent (red) and Flint (teal)
  parent lines.
  (\textbf{A}) Principal component analysis (PCA) of both
  groups for genomic (G), transcriptomic (T), leaf metabolic (L) and root
  metabolic (R) data. The variance explained by PC 1 (x axis) and PC 2 (y
  axis) are shown in the caption of each facet.
  (\textbf{B}) Linkage disequilibrium decay as a function of the distance
  between two loci using 40 bins of 0.125 Mbp width, each.
  The median $r^2$ is depicted as a horizontal bar whereas the mean $r^2$ is 
  depicted as a white diamond.
  }
\label{fig:PredictorProperties}
\end{figure}

We observed high median pairwise linkage disequilibrium (LD) between
SNP markers ($r^{2} \approx 0.39$ in Dent and $r^{2} \approx 0.37$ in Flint 
material) at a distance of $\Delta \text{Mbp} \leq 0.125$ 
(Fig. \ref{fig:PredictorProperties}\textit{B}). 
After an initial drop in $r^{2}$ for $\Delta > 0.125$, substantial long-range
LD remained.
Large differences in allele frequencies in the two heterotic groups were 
present for 57\% of SNPs (Fig. \ref{fig:Circos}A,B) --- particularly in the 
telomeric regions of the genome.
An eQTL analysis performed with the parent lines suggests that transcript 
abundance integrates variegated genetic information given the fact that i) on 
the same chromosome, significant associations not only occurred between 
adjacent but also between distant pairs of expressed genes and SNPs and ii) 50\% 
of the significant associations ($\alpha = 0.05$, Bonferroni-corrected) occurred
between expressed genes and SNPs on different chromosomes
(Fig. \ref{fig:Circos}).

% Figure 2
\begin{figure}[H]
\centering
\includegraphics[width=0.95\linewidth]{./figures/86_13_eQTL_Circos.pdf}
\caption{
  Distribution and relationship of genomic and transcriptomic
  data for the ten maize chromosomes.
  Centromeres for each chromosome are depicted as vertical red lines.
  (\textbf{A,C}) Density of SNPs and mRNAs, respectively, across the ten maize
  chromosomes.
  (\textbf{B,D}) Statistical significance ($-\log_{10} \text{p-value}$) for
  differences in SNP allele frequencies and expression levels of mRNAs,
  respectively, between Dent and Flint lines.
  (\textbf{Center}) Links between any statistically significant
  ($\alpha = 0.05$ after Bonferroni correction) association between SNPs and
  mRNAs.
  Associations are displayed as links for SNPs on chromosome 5, for which the
  distribution of associations is representative for the entire genome, using 
  red color for Dent parent lines and teal color for Flint parent lines.
}
\label{fig:Circos}
\end{figure}

\Genetics2level{Predictive abilities}
Assuming a polygenic architecture for all traits, as suggested by results from
a GWAS (Fig. S\ref{fig:GCA_GWAS_Manhattan}), we chose the best linear unbiased
predictor (BLUP) method as a baseline for prediction of T0, T1 and T2 hybrids.
Given that we corrected for fixed effects in advance, this method is equivalent
to a selection index.
A cross-validation scheme with 1,000 runs (CV1000), stratified by the parent 
lines, was devised (File S1, Fig. S\ref{fig:CV_Overview}).
Our main emphasis was on predicting T0 hybrids given the fact that they 
constitute the majority of possible hybrids in practical breeding programs 
\cite{Kadam2016}.

For predictive abilities ($r$) of T0 hybrids, transcriptomic data alone were the
best predictor ($\alpha = 0.01$) for the most complex and highly heterotic
trait, DMY, as well as for PRO (Fig. \ref{fig:PredAbility}\textit{A}).
Intriguingly, owing to the marked non-linear relationship between $r$ and
the probability of successfully selecting the best hybrid candidates
$P[r, \beta]$ (Fig. \ref{fig:PredAbility}\textit{B}), the 14.7\% higher $r$ for
DMY, when using transcriptomic over genomic data, yields an increase in
$P[r, \beta]$ of 85\% at a selection intensity $\beta = 0.01\%$.
This selection intensity corresponds to picking the top 100 out of $10^{6}$ 
predicted hybrids for production and intensive testing in field trials.

% Figure 3
\begin{figure*}[tbph]
\centering
\includegraphics[width=17.8cm,height=8.89cm]{./figures/t0_pred_and_success.pdf}
  \caption{
  Predictive abilities (r) from BLUP models using a CV scheme with 
  sampling of $|H_{TRN}| = 200$ hybrids, $|D_{TRN}|=40$ Dent and $|F_{TRN}|=33$ 
  Flint parent lines for various predictors and combinations thereof 
  (P, pedigree; G, genome; T, transcriptome; L, leaf metabolome; R, root 
  metabolome).
  (\textbf{A}) Comparison of $r$ values from 1,000 CV runs for T0 hybrids and
  seven agronomic traits.
  Differences among the means of predictive abilities were tested by Tukey's HSD 
  and the same letter was used for non-significant ($\alpha = 0.01$) 
  differences.
  (\textbf{B}) Success rate of selecting superior hybrids ($P[r,\beta]$).
  $P[r, \beta]$ is a function of the predictive ability $r=r(y,\hat{y})$ and
  refers to the conditional probability of a hybrid, selected at random from the 
  upper $\beta \text{\%}$ fraction of the distribution of predicted values 
  ($\hat{y}$), having a phenotypic value contained in the upper $\beta \text{\%}$ 
  of the distribution of phenotypic values $y$.
  Observed predictive abilities ($r$) for T0 hybrids and the trait DMY are 
  displayed as vertical, colored lines for three predictors.
  }
\label{fig:PredAbility}
\end{figure*}

Compared to other individual predictors, predictive abilities obtained with
genomic data alone were significantly ($\alpha = 0.01$) higher for FAT and SUG.
Root metabolites displayed moderate to high predictive abilities for DMY and
FAT, but did not perform well otherwise.
Leaf metabolites performed relatively poorly for all traits.
Regardless of the trait, combinations of genomic and transcriptomic information 
displayed robust and consistently high predictive abilities.
Except for PRO, incorporating additional endophenotypes as predictors into our 
models did not yield statistically significant ($\alpha = 0.01$) improvements 
but remained at the same level compared to combining genomic and transcriptomic 
data.
Incorporating SCA effects into our models did not further improve predictive
abilties ($\alpha = 0.05$, Fig. S\ref{fig:GCA_vs_SCA_PredAbility}).
Results for the combination of other predictors with metabolic data are not
presented because no improvement of predictive abilities over the combination 
of genomic data with transcriptomic data and pedigree data could be achieved.
Finally, we assessed the influence of the number of SNPs and mRNAs on predictive 
abilities. 
For genomic data, a subset of 5,000 SNPs already yielded the same predictive 
ability as when using the entire available set and for transcriptomic data, the
predictive ability improved only marginally with subsets larger than 50\% of 
the available transcripts (Fig. S\ref{fig:Feature_Filter}).

\section*{Discussion}
\Genetics2level{A paradigm shift in hybrid breeding}
Hybrid breeding programs are generally based on genetically divergent heterotic 
groups \cite{Bernardo2010}.
Their use enables a better exploitation of heterosis when conducting crosses 
between them \cite{Melchinger1998} and is expected to reduce the ratio of 
specific to general combining ability variance 
($\sigma^{2}_{\text{SCA}} : \sigma^{2}_{\text{GCA}}$) in the crosses, thereby 
allowing for the selection of hybrids largely on the basis of GCA of their 
parent lines \cite{Reif2007}.
However, obtaining accurate estimates of GCA requires the evaluation of 
new lines in combinations with testers from the opposite heterotic group in
multi-environment field trials.
The promise of hybrid prediction is to accelerate breeding programs by skipping 
a large share of these tests in favor of selecting the most promising hybrids
before they are even produced \cite{Technow2014}.
This approach involves the prediction of an impressive number of putative hybrid 
candidates ($n^{2}$) using predictor data collected on only 2$n$ parent lines.
Crucial for hybrid prediction are predictors, which not only reflect the 
relationship between parental inbred lines but also the interaction of the two 
parental genomes in their hybrid progeny.

\subsection{Heterotic groups}
Because of genetic drift and selection for hybrid performance, allele
frequencies are expected to diverge in the two heterotic groups, thereby
enlarging their genetic distance \cite{Falconer1996,Reif2007,Lariepe2017}.
Consistent with this hypothesis and two pilot studies with U.S. maize lines
\cite{Gerke2015,Hall2016}, Dent and Flint lines in our study were clearly 
separated in principal component analyses of genomic and transcriptomic data.
With large differences in allele frequencies $p^{I}$ and $p^{II}$ in the two
heterotic groups, as observed for 57\% of SNPs, dominance variance
$\sigma_{D}^{2}$ becomes very small because it is a function of the product
$p^{I} ( 1 - p^{I} ) p^{II} ( 1 - p^{II} )$ (\citeNP{Stuber1966},
File S1).
$\sigma_{D}^{2}$ is the main component contributing to the variance of 
the specific combining ability  effects ($\sigma_{\text{SCA}}^{2}$), describing 
interactions among the parental genomes in hybrid combinations.
It was therefore not surprising that the variances of the general combining
ability (GCA) effects ($\sigma_{\text{GCA}^{D}}^{2}$ and 
$\sigma_{\text{GCA}^{F}}^{2}$), describing the main effects of the parents from 
each heterotic group, together explained more than 93\% of the genotypic 
variance among hybrids for agronomic traits, which is consistent with an earlier
study on silage maize of the Dent $\times$ Flint heterotic pattern
\cite{Argillier2000}.
The importance of GCA in our material was further corroborated by statistically 
non-significant ($\alpha = 0.05$) differences in predictive abilities between 
models using only GCA effects and those that additionally incorporated SCA 
effects (Fig. S\ref{fig:GCA_vs_SCA_PredAbility}).
Nevertheless, in crops such as wheat, with no clearly defined heterotic groups
\cite{Zhao2015a} and greater importance of SCA, inclusion of SCA effects in
the model should improve predictive abilities.

\subsection{Properties of well-established predictors}
While pedigree data reflect the expected relationship between individuals, they 
do not necessarily depict their realized relationship.
Genomic data and downstream endophenotypes offer to improve upon this
pedigree-based approximation by more closely mirroring the transmission of 
genes between individuals and their interactions.
Genomic data have the advantage of reliably capturing Mendelian sampling, 
thereby improving pedigree-based prediction for many traits.
However, genomic data alone may not be the final answer for the prediction of
complex traits for two major reasons:
First, the number of samples in most studies is considerably smaller than the
number of genetic markers or even nucleotides of a genome.
This implies that just modeling additive effects already necessitates effect 
shrinkage.
More importantly, however, interactions between loci throughout the genome 
can be frequent \cite{Brem2005,Brown2014}, but attempts to incorporate this 
epistasis for the prediction of heterotic traits using genomic data have been 
disappointing when the prediction and the training set did not share the same
or closely related parents \cite{Jiang2015}.
This was true even when using recently developed, efficient models 
\cite{Jarquin2014,Martini2016} and suggests that genomic data capture 
only statistical epistasis, referring to genetic variation at the population 
level \cite{Sackton2016}, which is generally of negligible magnitude 
\cite{Hill2008,Mackay2014,Guo2016,Vazquez2016}.

\Genetics2level{Complementation of predictors}
\subsection{Flow of biological information}
It is well-known that genetic effects on the phenotype are mediated through
multiple layers of endophenotypes \cite{Civelek2014,Ritchie2015b} with 
information mainly flowing from the genome toward the phenotype via the 
transcriptome, the proteome and the metabolome with metabolite fluxes 
ultimately governing energy production and growth \cite{Cornish-Bowden2001}.
For most traits in our material, metabolite- and pedigree-based predictive
abilities were lower than those obtained with either transcriptomic or genomic
information.
However, consistently high predictive abilities across multiple traits could be
realized when combining multiple predictors, as has been reported previously in 
humans \cite{Vazquez2016} and maize inbred lines \cite{Guo2016}.
This suggests complementary properties of the different predictors resulting in 
better proxies for the complex interplay in gene networks than solely genomic 
information.
Such an advantage is particularly important for hybrid prediction when parents 
of prediction set hybrids are not closely related to parents of training set
hybrids (File S1) as was shown by the relative excellence of transcriptomic 
data and the use of multiple predictors for the prediction of traits in T0 
hybrids compared to T1 and T2 hybrids.

\subsection{Tapping new sources of information}
Whereas pedigree and genomic information are static, subsequent endophenotypes 
are characterized by pervasive interactions among and between each other 
\cite{Dalchau2011,Zhu2012} and are, to varying degrees, influenced by biotic 
\cite{Rudd2015,Tzin2015} and abiotic perturbations \cite{Caldana2011,Witt2012}.
So while endophenotypes do not exclusively report on physiological epistasis but 
also on non-heritable effects, they do seem to capture important information not
represented by the genome given their intermediate position in the
genotype-phenotype cascade.
We get support for this hypothesis from (i) merely low to moderate correlations 
between off-diagonal elements of the kernels of different predictors in our 
study, (ii) mounting evidence for further improvements of predictive abilities 
when complementing genomic prediction with other endophenotypes despite 
sufficient marker densities (Fig. S\ref{fig:Feature_Filter}, \citeNP{Guo2016})
and (iii) the integration of SNP information from close and distant eQTL in the 
transcripts analyzed in our study.
However, we concede that the number of parental genotypes in our mRNA assays
was too small to warrant a reasonable statistical power for detecting epistasis
in the expression of transcripts.
While two other studies also attempted to model interactions between different
predictors, we refrained from this approach given that their reported predictive
abilities based on interactions were not different from those in additive
models despite using much larger sample sizes \cite{Vazquez2016,Guo2016}.

\Genetics2level{Transcriptomic data}
\subsection{Utility of transcriptomic data for trait predictions}
Of particular note was the excellent performance of transcriptomic data in 
predicting dry matter yield and protein.
Evidence that parental gene expression patterns might be predictive of hybrid 
performance is given by (i) prevailing additive expression patterns in maize 
hybrids \cite{Springer2007,Stupar2008}, (ii) a positive correlation of the 
proportion of additive gene expression with the yield of hybrids \cite{Guo2006}, 
and (iii) co-localization of additively expressed genes with heterotic QTL 
\cite{Thiemann2014}.
According to metabolic flux theory, gene expression in hybrids at the 
mid-parent level can generate hybrid vigor by counterbalancing opposing 
detrimental expression levels in their parent lines on a genome-wide scale 
\cite{Kacser1981,Springer2007a}.
The same concept is expected to apply to other quantitative endophenotypes 
\cite{Lisec2011}.

\subsection{Pre-selection bias}
As pointed out earlier, our transcripts were pre-selected based on associations
with grain dry matter yield and grain dry matter content in hybrids, using a 
subset of the data included in our study ($H_{\text{FAC}^{(1)}}$).
Two findings indicate that no pre-selection was introduced:
(i) Relative differences in predictive abilities between transcriptomic and
pedigree or genomic data did not change when excluding genotypes from 
$H_{\text{FAC}^{(1)}}$ from the data (Fig. S\ref{fig:NoExp1}) and 
(ii) transcriptomic data performed rather poorly in predicting dry matter 
content although this trait was also among the criteria for the pre-selection 
procedure.
Finally, an independent study using RNA-Seq data for the prediction of traits
in maize inbred lines also reported exceptionally good performance of 
transcriptomic data in the prediction of multiple yield-related traits 
\cite{Guo2016}.

\Genetics2level{Relative excellence of predictors for different traits}
\subsection{Tissue and sampling time}
Despite the great prospects of using endophenotypes for trait predictions, some
aspects require careful consideration when using this approach.
A particular challenge in endophenotype-based prediction efforts is the selection
of a suitable tissue and sampling time.
Tissue-related effects regarding gene expression were found in studies on humans 
\cite{Yang2015,Mele2015,Searle2016} and \textit{A. thaliana} \cite{Schmid2005}
and in maize hybrids with respect to metabolome composition and metabolite 
abundance \cite{Witt2012}.
Moreover, the age of an organism can selectively influence the expression of 
genes as observed in studies on humans \cite{Mele2015,Yang2015} and 
\textit{C. elegans} \cite{Vinuela2010,Francesconi2014}.
The low correlations between the off-diagonal elements of the kernels calculated 
from root and leaf metabolites might therefore be a reflection of highly dynamic 
processes differing between tissues and during different developmental stages.
Whereas root metabolic data and transcriptomic data were obtained from
seedlings germinated in standard controlled conditions, leaf metabolic data were
derived from field-grown plants at a much later developmental stage, thereby
increasing the possibility of environmentally-induced modifications.
One might hypothesize, that the choice of sampling time and tissue could
influence the chances of successful trait prediction if such age- or 
tissue-dependent transcripts and metabolites are associated with a phenotypic 
or clinical trait.
Generally, sampling from the seedling would be highly desirable in breeding when 
trying to scale the use of endophenotypic data to even larger numbers of inbred 
lines and assay genotypes independent of the season.

\subsection{Feature selection}
Another explanation for trait-dependent excellence of any predictor might lie
in the sampling of features.
In this study, only a small subset of metabolites was sampled and even very 
recent technologies \cite{Xu2016,Dan2016} capture only a fraction of the 
estimated set of metabolites \cite{Fernie2007}.
Moreover, the smaller differences in metabolite levels between both heterotic 
groups (Fig. \ref{fig:PredictorProperties}) were most likely not conducive to
capturing basic components underlying complex heterotic traits.
It is also possible that transcriptomic data are associated with more
biological processes than metabolite data and better capture the genetic effects
relevant for the prediction of T0 hybrids.

\subsection{Prospects for metabolites}
Previously observed moderate metabolite-based predictive abilities for T1
hybrids \cite{Riedelsheimer2012b} were confirmed in our study
(Fig. S\ref{fig:PredAbilityT0-T2}), but for the majority of traits, root 
metabolites reached only medium and leaf metabolites even lower predictive 
abilities when predicting T0 hybrids.
Despite the aforementioned shortcomings of metabolites, they have shown to be 
intriguing predictors due to their phsyiological proximity to the phenotype,
which provides information that is impossible to infer from DNA or proteins
\cite{Fernie2012}, as well as encouraging results from other studies
\cite{Guo2016,Dan2016}.
A recently introduced technology, allowing for live-measurements of small 
molecules in the blood of living and awake animals \cite{Arroyo-Curras2017}, 
might overcome the problem of poorly time-resolved snap-shots of some 
metabolites with extremely fast turnover rates \cite{Arrivault2009} if modified
to properly work in plants.

\section*{Conclusions}
The use of whole-genome information has considerably advanced trait prediction
over traditional pedigree-based BLUP by incorporating previously unobservable 
Mendelian sampling.
Combining variegated sources of information promises to capture complex
interactions between genes and endophenotypes, leading to stable predictions
across traits.
Especially if an extremely small fraction of the candidates is selected from
the millions of possible new hybrids from each breeding cycle, the success of
forecasts is a strongly convex function of predictive ability
(Fig. \ref{fig:PredAbility}\textit{B}).
Therefore, considering endophenotypes could have a substantial effect on the 
success and economics of hybrid breeding.
Given the anticipated technological improvements in RNA-Seq and metabolite 
profiling, as well as the forthcoming adoption of the DH-technology for many
crops \cite{Kelliher2017}, a paradigm shift from exclusively genomic prediction 
models to more inclusive approaches seems imminent.

\section{Acknowledgments} 
We thank the staff of the Agricultural Experimental Research station,
University of Hohenheim, for excellent technical assistance in
conducting the field experiments. We are indebted to the group of R.
Fries from Technische Universit{\"a}t M{\"u}nchen for the SNP genotyping of the
parent inbred lines, to X. Mi for his assistance in preparing auxiliary
figures based on the Mathematica software, to C. Zenke for advice on the
computation of transcriptomic BLUEs and to P. Schopp for advice on
prediction models. The authors acknowledge support by the state of
Baden-W{\"u}rttemberg through bwHPC. This project was funded by the German
Federal Ministry of Education and Research (BMBF) within the projects
OPTIMAL (FKZ: 0315958B,0315958F), SYNBREED (FKZ: 0315528D) and by the
German Research Foundation (DFG, Grants No. ME 2260/5-1 and SCHO 764/6-1).
Financial support for M.W. was provided by the Fiat Panis foundation, Ulm,
Germany.

\nolinenumbers
% Bibliography
\bibliography{library}
\bibliographystyle{mychicago}
\end{document}