### 2017-06-15 reports/select_subpopulation.Rmd
Prepare random subsampling for prediction

---
title: Selection of a subpopulation with little structure
author: Matthias Westhues
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
    fig_caption: yes
bibliography: ["~/Dropbox/Paper/library.bib"]
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
# Data and packages -------------------------------------------------------
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse", "LEA", "stringr", "viridis", "cowplot", 
               "ggthemes", "pophelper", "corehunter", "forcats")
#devtools::install_github("mwesthues/sspredr")
pacman::p_load_gh("mwesthues/sspredr")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  fig.width = 8, fig.height = 6, warning = FALSE, echo = FALSE, 
  message = FALSE
  )
```

# Rationale {-}
Based on our discussion with G. Thaller, R.L. Fernando and C. Heuer, we 
concluded that the pre-selection of a subset of individuals might be necessary
to mitigate the impact of population structure on predictive ability.

```{r pca-structure, cache=TRUE, results='hide'}
# For the second scenario, keep only names of genotypes, which are covered by 
# all data types (i.e. phenotypic, genotypic and transcriptomic).
common_genotypes <- readRDS(
  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
)
common_genotypes <- common_genotypes %>%
  reduce(intersect)

snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
  readRDS() %>%
  #.[rownames(.) %in% common_genotypes, ] %>%
  sspredr::ensure_snp_quality(
    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05, 
    any_missing = FALSE, remove_duplicated = TRUE
  )

# PCA ---------------------------------------------------------------------
write.lfmm(snp, "./data/derived/maizego/snp.lfmm")

# Determine the structure of the data using genotypic data.
snp_pc <- LEA::pca(input.file = "./data/derived/maizego/snp.lfmm",
                   scale = TRUE)

pc_mat <- snp_pc$projections
rownames(pc_mat) <- rownames(snp)
colnames(pc_mat) <- paste0("PC_", seq_len(ncol(pc_mat)))

pc_df <- pc_mat %>%
  as.data.frame() %>%
  rownames_to_column(var = "G") %>%
  gather(key = PC, value = Score, -G) %>%
  as_data_frame() %>%
  filter(PC %in% paste0("PC_", seq_len(2))) %>%
  mutate(
    PC = gsub("PC_", replacement = "PC", x = PC)
  ) %>%
  spread(key = PC, value = Score)

# PCA ---------------------------------------------------------------------
genos <- read_tsv(
  "./data/processed/maizego/popstruc.names",
  col_names = FALSE
  ) %>% 
  select(2) %>% 
  flatten_chr()

# Select K=4
# For each individual, return the cluster with the largest coefficient.
# Assign a color to each cluster.
# Plot the PCA results with colors corresponding to the cluster with the largest
# coefficient for that individual.

get_max_cluster <- function(x) {
  x %>%   
    mutate(G = genos) %>% 
    gather(key = "Cluster", value = "AncCoef", -G) %>% 
    group_by(G) %>% 
    summarize(main_cluster = which.max(AncCoef)) %>% 
    mutate(main_cluster = main_cluster %>% as.character() %>% as.factor()) %>% 
    ungroup()
}

# STRUCTURE FILES ---------------------------------------------------------
slist <- list.files(
  path = "./data/processed/maizego",
  pattern = "popstruc[0-9]",
  full.names = TRUE
  ) %>% 
  readQ(
    files = .,
    indlabfromfile = TRUE,
    filetype = "structure"
  )

cluster_df <- slist %>% 
  map(get_max_cluster) %>% 
  map(.f = ~right_join(., y = pc_df, by = "G")) %>% 
  bind_rows(.id = "K")

change_to_k <- function(x) {
  x %>% 
    gsub(x = ., replacement = "", pattern = "_f") %>% 
    gsub(x = ., replacement = "", pattern = "popstruc")
}
```

# Analysis {-}
First, I ran STRUCTURE [@Pritchard2000] to estimate the ancestry coefficients
for each of the `r nrow(snp)` genotypes from the subset 'tropical/subtropical' 
of the [maize diversity panel](http://www.maizego.org/Resources.html).
In total, I have examined 8 putative ancestral populations (K=2 to K=9) and, 
based on combinations of bar plots and PCA plots 
(Fig. \@ref(fig:pca-structure-plot)), as well as the estimated probability of 
the data for each K, I concluded that K=4 represents the population structure 
best (Fig. \@ref(fig:pca-structure-plot)).

```{r pca-structure-plot, fig.height=12, fig.cap="Population structure of a set of 211 inbred lines from the tropical/subtropical groups of the maize diversity panel. ($\\textbf{A}$) Structure coefficients assuming K=4 ancestral populations. ($\\textbf{B}$) Principal Component Analysis. Individual points are colored according to the cluster with which they have the highest overlap."}
k <- 4
g1 <- slist %>% 
  set_names(., change_to_k(names(.))) %>% 
  map(.f = ~mutate(., G = genos)) %>% 
  keep(names(.) == k) %>% 
  map(.f = ~mutate(., Central_Cluster = Cluster1 + Cluster4)) %>% 
  map(.f = ~gather(
    ., key = "Cluster", value = "AncCoef", -G, -Central_Cluster)
    ) %>% 
  bind_rows(.id = "K") %>% 
  mutate(G = as.factor(G)) %>% 
  ggplot(aes(
    x = fct_reorder(G, x = Central_Cluster), y = AncCoef, fill = Cluster)
    ) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_blank(),
    legend.position = "top"
  ) +
  guides(fill = guide_legend(ncol = 2)) +
  xlab("Genotype") +
  ylab("Ancestry Coefficient")

g2 <- cluster_df %>%
  mutate(K = change_to_k(K)) %>% 
  filter(K == k) %>% 
  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
  geom_point(size = 1) +
  guides(color = guide_legend(
    override.aes = list(size = 3),
    title = "Cluster"
    )) +
  scale_color_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside"
  )

plot_grid(
  g1, g2, labels = c("A", "B"), nrow = 2
)
```

The set of selected individuals (Fig. \@ref(fig:pca-plot)) roughly corresponds
to clusters 1 and 4 in Fig. \@ref(fig:pca-structure-plot) and excludes 
individuals in the tails of the PCA plot.

```{r pca-plot, fig.cap="Selecting individuals based on PCA"}
selected_df <- slist %>% 
  set_names(., change_to_k(names(.))) %>% 
  map(.f = ~mutate(., G = genos)) %>% 
  keep(names(.) == k) %>% 
  .[[1]] %>% 
  left_join(y = pc_df, by = "G") %>% 
  mutate(main_cluster = case_when(
    PC1 >= -30 & PC1 <= 10 & PC2 >= -25 & PC2 <= 40 ~ "yes",
    PC1 != 0 ~ "no"
    ))

selected_df %>%
  mutate(main_cluster = as.factor(main_cluster))  %>% 
  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
  geom_point(size = 1) +
  guides(color = guide_legend(
    override.aes = list(size = 3),
    title = "Selected"
    )) +
  scale_color_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside"
  )
```

```{r sub-pca, cache=TRUE, results='hide', fig.cap="PCA with all 211 inbred lines from the tropical/subtropical subset of the maize diversity panel. Individuals with seemingly negligible structure are being selected (blue dots)."}
selected_geno <- selected_df %>% 
  filter(main_cluster == "yes") %>% 
  select(G) %>% 
  flatten_chr()

sel_snps <- snp %>% 
  .[rownames(snp) %in% selected_geno, ] %>% 
  sspredr::ensure_snp_quality(
    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05, 
    any_missing = FALSE, remove_duplicated = TRUE
  )
write.lfmm(sel_snps, "./data/derived/maizego/sel_snps.lfmm")
# Determine the structure of the data using genotypic data.
sel_snp_pc <- LEA::pca(
  input.file = "./data/derived/maizego/sel_snps.lfmm",
  scale = TRUE
  )

sel_pc_mat <- sel_snp_pc$projections
rownames(sel_pc_mat) <- rownames(sel_snps)
colnames(sel_pc_mat) <- paste0("PC_", seq_len(ncol(sel_pc_mat)))
```

```{r sub-pca-plot, fig.cap="PCA of the selcted inviduals from Fig. 2."}
sel_pc_df <- sel_pc_mat %>%
  as.data.frame() %>%
  rownames_to_column(var = "G") %>%
  gather(key = PC, value = Score, -G) %>%
  as_data_frame() %>%
  filter(PC %in% paste0("PC_", seq_len(2))) %>%
  mutate(
    PC = gsub("PC_", replacement = "PC", x = PC)
  ) %>%
  spread(key = PC, value = Score)

sel_pc_df %>% 
  ggplot(aes(PC1, PC2)) +
  geom_point() +
  theme_pander(base_size = 15)
```

Finally, I ran a PCA on the subset of `r nrow(sel_snps)` selected individuals 
to check whether there is still structure in the population 
(Fig. \@ref(fig:sub-pca-plot)).
I'm not pleased with this result as this plot still suggests structure in the
subpopulation.
We might argue, as G. Thaller suggested, that a breeding population will always 
be structured to an extent, however, I doubt that a reviewer would be convinced
by our rationale of subsetting a population when the subset itself still 
displays this structure.

## Possible solution {-}
Should we omit Fig. \@ref(fig:sub-pca-plot) and present only 
Fig. \@ref(fig:pca-plot)?

# References