### 2017-09-01 reports/resampling_report.Rmd
init resampling pred report

---
title: Selection of a subpopulation with little structure
author: Matthias Westhues
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: readable
    fig_caption: yes
highlight: tango
documentclass: book
bibliography: ["~/Dropbox/Paper/library.bib"]
biblio-style: apalike
link-citations: yes
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
# Data and packages -------------------------------------------------------
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  "tidyverse",
  "viridis",
  "ggthemes",
  "forcats",
  "LEA",
  "pophelper",
  "cowplot",
  "dtplyr",
  "data.table",
  "gsubfn"
  )
#devtools::install_github("mwesthues/sspredr")
pacman::p_load_gh("mwesthues/sspredr")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  fig.width = 8, fig.height = 6, warning = FALSE, echo = FALSE,
  message = FALSE
  )
```

# Overview
The purpose of this project is to explore the utility of single-step prediction
of phenotype performance in a set of maize inbred lines and maize hybrids,
respectively.
The methods used herein are based on the seminal papers by @Legarra2009 and
@Christensen2010 with extensions introduced by @Fernando2014.

# Data Preparation
We analyze hybrid maize data from the public breeding program of the University
of Hohenheim as well as [data on inbred lines](http://www.maizego.org/Resources.html) from the lab of doctor Jianbing Yan.
The data on inbred lines were downloaded on 2017-01-31 from the [Baidu Cloud](https://pan.baidu.com/s/1eQH3hfW#list/path=%2F)
set up by the Yan lab and are stored under `data/input/maizego`.
The Hohenheim data were generated as part of the publication "Omics-based Hybrid
Prediction in Maize" [@Westhues2017].

## Agronomic Data
For the maize inbred lines we focused on the following traits that were
previously described in [Guo et al. (2016)](https://link.springer.com/article/10.1007/s00122-016-2780-5):

*   `Cob weight` (CW)
*   `Days to silking` (DS)
*   `Ear diameter` (ED)
*   `100-grain weight` (GW)
*   `Kernel width` (KW)
*   `Plant height` (PH)

and explored the explored the distribution (including skewness) of each trait
and the phenotypic correlation between traits.

In the case of maize hybrids we analyzed the traits

*   `Dry matter yield` (DMY)
*   `Dry matter content` (DMC)
*   `Fat` (FAT)
*   `Protein` (PRO)
*   `Starch` (STA)
*   `Sugar` (SUG)

## Genomic Data
The preparation of genomic marker data for hybrids and inbred lines,
respectively, involved the following five steps (for hybrid data these were
applied separately to each of the two heterotic groups):

1.   Remove loci with a callfrequency $\geq 0.95$.

2.   Remove loci with heterozygosity $\geq 0.05$.

3.   Remove loci with minor allele frequency $\geq 0.05$.

4.   Impute remaining missing values using BEAGLE using 25 iterations and 20
     samples per iterations.

5.   Remove copies (i.e. loci in perfect LD with a previous one) of marker
     loci.

## Transcriptomic Data
For both material types transcriptomic data were already sufficiently
pre-processed.

## Population Structure
Principal component analyses were run for the [hybrid](analysis/pca.R) and
[inbred](analysis/maizego_snp_analyses.R) material.

The separation between the Dent and the Flint heterotic group for the hybrid
material [looked as expected](analysis/pca.R) but for the inbred lines we
found unexpected clustering of the lines, as determined via a [STRUCTURE](analysis/maizego_structure.R),
analysis and a [PCA](analysis/maizego_snp_anlyses.R), albeit it with very low
fractions of variance explained by the first five principal components.
See \@ref(population-structure) for a more detailed report.
Inbred lines that belonged to the clusters `1` and `4`, which were rather
homogeneous, were assigned to scenario `A`.
We defined a second scenario `B` which simply comprised all available inbred
lines and ran it to have predictive abilities for all genotypes in case
population structure was no issue for this material.

# Predictions
## Preparation
### Previous work for this manuscript
In the first draft we wanted to evaluate the influence of the genetic
constitutation of the set of genotypes that has only data on one out of two
predictors by generating core samples  (https://github.com/mwesthues/single_step/commit/8f8b39198be7d5091954f84ec0c5834afd1c3cfa).
We defined core samples as a subset of genotypes that is covered by two
predictors whereas all other genotypes (*i.e.* the complement) are only covered
by a single predictor.
The first application involved assembling core sets of varying sizes by
maximizing the average genetic distance among core set members using the
Modifed Rogers (MR) distance as the criterion ([Thachuk et al. (2009)](https://bmcbioinformatics.biomedcentral.com/articles/10.1186/1471-2105-10-243)).
The size of the core set was varied in increments of ten percentage points and
ranged from 10% to 90% of all inbred lines.
The issue with this approch is that we could not preclude effects of population
structure in our material on predictive abilities.
Therefore, we decided to replace the core sampling procedure by a nested random
subsampling procedure.

### Nested subsampling scheme
To ensure that a potential bias due to population structure would be
represented by an increased standard error of the predictive abilities we
conceived a nested resampling scheme.
A first level of sampling was applied to each combination of `Material` (*i.e.*
"Hybrid" or "Inbred"), `Scenario` ("A" and "B" for inbred lines; "None" for
hybrids), `Extent` ("Core" if every predictor was available for all genotypes
and "Full for the entire set of genotypes") and `Core_Fraction`, representing
the share of core-set genotypes for which data on all predictors were assumed
to be available.
The notion of `Core_Fraction` may sound contradictory to the concept of "Core"
and "Full" sets of genotypes but note that for the variable `Core_Fraction` we
artificially removed information on a predictor for some genotypes to implore
the influence of genetic space-coverage through incomplete predictors on
predictive ability.
This random subsampling procedure was repeated 20 times and stored as
`Rnd_Level2`.

For all core-set inbred lines we applied a second, nested randomization scheme
where `n * frac` genotypes were declared to have both data on the complete and
the incomplete predictor whereas `n - n * frac` genotypes had only information
on the complete predictor.
Here `frac` denotes the fraction of all genotypes having information on both
predictors.
This second randomization was nested within each randomization stored in
`Rnd_Level2` and applied 20 times per randomization in `Rnd_Level2`.
Therefore, we ended up with 20 sets of predictions for each combination of
`"Material" * "Extent" * "Scenario"` when `frac = 1` and `20 ** 2 = 400`
sets of predictions for each combination of `"Material" * "Extent" *
"Scenario" * "Core_Fraction"` when `frac != 1`.

### Prediction template
In order to use the resources on the server as efficiently as possible,
similar combinations of the various parameters considered (`Trait`,
`Predictor`, `Material`, `Extent`, `Scenario`, `Rnd_Level1`, `Rnd_Level2`,
`Core_Fraction`) were clustered together (coded as `Interval`) in a
prediction template.

After running some speed tests for some of these combinations a
script for the automatic generation of `msub` commands was built.

## Execution
All predictions were run on the [bwunicluster](https://www.bwhpc-c5.de/wiki/index.php/Category:BwUniCluster)
by entering the msub commands into the terminal on the cluster.
For information on how to use the `bwunicluster`, in case you have access to
it, can be found [here](https://mwesthues.github.io/bwunicluster.html).

After running the prediction, the results from all clusters were concatenated.

## Bootstrap
Separately for each combination of `Material`, `Extent`, `Scenario`, `Trait`,
`Core_Fraction`, `Predictor`, `Rnd_Level1` and `Rnd_Level2`, a
bootstrap of 10,000 runs was applied to a data frame comprising observed and
predicted phenotypic values.
This yielded 10,000 bootstrap predictive abilities for each of the
aforementioned combinations together with a corresponding standard error.
The final predictive ability was calculated by taking the arithmetic mean
of bootstrapped predictive abilities over `Rnd_Level1` and `Rnd_Level2`.
The corresponding standard error was calculated across the average bootstrapped
standard error across `Rnd_Level1` and `Rnd_Level`.

## Visualization
```{r prediction-data, cache=FALSE, results='hide'}
boot_df <- "./data/processed/predictions/bootstrapped_predictions.RDS" %>%
  readRDS() %>%
  dplyr::group_by(Combi, Trait, Core_Fraction, Predictor) %>%
  dplyr::summarize(
    avg_r = mean(r),
    se = mean(std.error)
  ) %>%
  dplyr::ungroup()

## -- HYBRIDS -----------------------------------------------------------------
# Specify a unified predictor order.
pred_order <- c("P", "G", "T", "PG", "PT", "GT")

# Specify a unique trait order.
trait_order <- c("DMY", "DMC", "ADL", "FAT", "PRO", "STA", "SUG")

hyb_df <- boot_df %>%
  dplyr::filter(Combi %in% c("CHN", "FHN")) %>%
  dplyr::mutate(Trait = forcats::fct_recode(
    Trait,
    "DMY" = "GTM",
    "DMC" = "GTS",
    "FAT" = "FETT",
    "PRO" = "RPR",
    "SUG" = "XZ"
  )) %>%
  dplyr::mutate(Combi = forcats::fct_recode(
    Combi,
    "Core" = "CHN",
    "Full" = "FHN"
  )) %>%
  dplyr::mutate(Trait = forcats::fct_relevel(Trait, trait_order)) %>%
  dplyr::mutate(
    Predictor = gsubfn::gsubfn(
      pattern = "\\S+",
      replacement = list(
        "mrna" = "T",
        "ped" = "P",
        "snp" = "G",
        "ped_snp" = "PG",
        "ped_mrna" = "PT",
        "snp_mrna" = "GT"
      ),
      x = Predictor
    ),
    Predictor = forcats::fct_relevel(Predictor, pred_order)
  )

# Because the bars and errorbars have different widths
# we need to specify how wide the objects we are dodging are
dodge <- position_dodge(width = 0.9)

# Predefine the color scheme.
mycols <- scales::brewer_pal(type = "div", palette = "Spectral")(n = 6)
names(mycols) <- c("P", "G", "T", "PG", "PT", "GT")
limits <- aes(
  ymax = `Predictive Ability` + se,
  ymin = `Predictive Ability` - se
  )

## -- INBREDS CORE_FRACTION == 1 ----------------------------------------------
inbred_df <- boot_df %>%
  dplyr::filter(!Combi %in% c("CHN", "FHN")) %>%
  dplyr::mutate(
    Predictor = gsubfn::gsubfn(
      pattern = "\\S+",
      replacement = list(
        "snp" = "G",
        "mrna" = "T",
        "snp_mrna" = "GT"
      ),
      x = Predictor
    ),
    Trait = gsubfn::gsubfn(
      pattern = "\\S+",
      replacement = list(
        "Plantheight" = "PH",
        "Eardiameter" = "ED",
        "100grainweight" = "GW",
        "cobweight" = "CW",
        "Kernelwidth" = "KW",
        "Silkingtime" = "DS"
      ),
      x = Trait
    )
  ) %>%
  dplyr::mutate(Combi = forcats::fct_recode(
    Combi,
    "Core" = "CHN",
    "Full" = "FHN"
  ))

# Predefine the color scheme.
# Use also the levels of predictors in the hybrid data set to ensure that,
# regardless of the material, the same color is picked for the same predictor.
a_colors <- scales::brewer_pal(type = "div", palette = "Spectral")(n = 6) %>%
  set_names(c("P", "G", "T", "PG", "PT", "GT")) %>%
  .[names(.) %in% c("G", "T", "GT")]
```

```{r hybrid-plot, cache=FALSE, fig.cap="Hybrid Plot"}
hyb_df %>%
  dplyr::rename(`Predictive Ability` = "avg_r") %>%
  ggplot(aes(x = Trait, y = `Predictive Ability`, fill = Predictor)) +
  geom_bar(stat = "identity", position = dodge, color = "black") +
  geom_errorbar(limits, position = dodge, width = 0.25) +
  facet_grid(. ~ Combi) +
  scale_fill_manual(values = mycols) +
  ggthemes::theme_pander(base_size = 10)
```

```{r inbred-plot, cache=FALSE, fig.cap="Inbred Plot"}
inbred_df %>%
  dplyr::filter(Core_Fraction == "1") %>%
  dplyr::rename(`Predictive Ability` = "avg_r") %>%
  ggplot(aes(x = Trait, y = `Predictive Ability`, fill = Predictor)) +
  geom_bar(stat = "identity", position = dodge) +
  geom_errorbar(limits, position = dodge, width = 0.25) +
  facet_grid(Combi ~ .) +
  scale_fill_manual(values = a_colors) +
  theme_pander(base_size = 10)
```

```{r core-inbred-plot, cache=FALSE, fig.cap="Core Plot"}
inbred_df %>%
  dplyr::filter(Predictor == "GT", Combi %in% c("CIA", "CIB", "CIC")) %>%
  dplyr::mutate_at(vars(Predictor), funs(as.factor)) %>%
  dplyr::rename(`Core Fraction` = "Core_Fraction") %>%
  dplyr::rename(`Predictive Ability` = "avg_r") %>%
  ggplot(aes(
    x = `Core Fraction`,
    y = `Predictive Ability`,
    color = Trait,
    group = Trait
    )
  ) +
  geom_line(stat = "identity", position = position_dodge(width = 0.3)) +
  facet_grid(~ Combi) +
  ggthemes::scale_color_tableau() +
  ggthemes::theme_pander(base_size = 10) +
  theme(legend.position = "right") +
  guides(color = guide_legend(override.aes = list(size = 3)))
```

# Appendix {-}
## Population Structure {-}
Based on our last discussion we concluded that the pre-selection of a subset
of individuals might be necessary to mitigate the impact of population structure
on predictive ability.

```{r pca-structure, cache=TRUE, results='hide'}
# For the second scenario, keep only names of genotypes, which are covered by
# all data types (i.e. phenotypic, genotypic and transcriptomic).
common_genotypes <- readRDS(
  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
)
common_genotypes <- common_genotypes %>%
  purrr::reduce(intersect)

snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
  readRDS() %>%
  #.[rownames(.) %in% common_genotypes, ] %>%
  sspredr::ensure_snp_quality(
    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05,
    any_missing = FALSE, remove_duplicated = TRUE
  )

# PCA ---------------------------------------------------------------------
write.lfmm(snp, "./data/derived/maizego/snp.lfmm")

# Determine the structure of the data using genotypic data.
snp_pc <- LEA::pca(input.file = "./data/derived/maizego/snp.lfmm",
                   scale = TRUE)

pc_mat <- snp_pc$projections
rownames(pc_mat) <- rownames(snp)
colnames(pc_mat) <- paste0("PC_", seq_len(ncol(pc_mat)))

pc_df <- pc_mat %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "G") %>%
  tidyr::gather(key = PC, value = Score, -G) %>%
  tibble::as_data_frame() %>%
  dplyr::filter(PC %in% paste0("PC_", seq_len(2))) %>%
  dplyr::mutate(
    PC = gsub("PC_", replacement = "PC", x = PC)
  ) %>%
  tidyr::spread(key = PC, value = Score)

# PCA ---------------------------------------------------------------------
genos <- readr::read_tsv(
  "./data/processed/maizego/popstruc.names",
  col_names = FALSE
  ) %>%
  dplyr::select(2) %>%
  purrr::flatten_chr()

# Select K=4
# For each individual, return the cluster with the largest coefficient.
# Assign a color to each cluster.
# Plot the PCA results with colors corresponding to the cluster with the largest
# coefficient for that individual.

get_max_cluster <- function(x) {
  x %>%
    dplyr::mutate(G = genos) %>%
    tidyr::gather(key = "Cluster", value = "AncCoef", -G) %>%
    dplyr::group_by(G) %>%
    dplyr::summarize(main_cluster = which.max(AncCoef)) %>%
    dplyr::mutate(main_cluster = main_cluster %>%
      as.character() %>%
      as.factor()
    ) %>%
    dplyr::ungroup()
}

# STRUCTURE FILES ---------------------------------------------------------
slist <- list.files(
  path = "./data/processed/maizego",
  pattern = "popstruc[0-9]",
  full.names = TRUE
  ) %>%
  pophelper::readQ(
    files = .,
    indlabfromfile = TRUE,
    filetype = "structure"
  )

cluster_df <- slist %>%
  purrr::map(get_max_cluster) %>%
  purrr::map(.f = ~dplyr::right_join(., y = pc_df, by = "G")) %>%
  dplyr::bind_rows(.id = "K")

change_to_k <- function(x) {
  x %>%
    gsub(x = ., replacement = "", pattern = "_f") %>%
    gsub(x = ., replacement = "", pattern = "popstruc")
}
```

First, I ran STRUCTURE [@Pritchard2000] to estimate the ancestry coefficients
for each of the `r nrow(snp)` genotypes from the subset 'tropical/subtropical'
of the [maize diversity panel](http://www.maizego.org/Resources.html).
In total, I have examined 8 putative ancestral populations (K=2 to K=9) and,
based on combinations of bar plots and PCA plots
(Fig. \@ref(fig:pca-structure-plot)), as well as the estimated probability of
the data for each K, I concluded that K=4 represents the population structure
best (Fig. \@ref(fig:pca-structure-plot)).

```{r pca-structure-plot, fig.height=12, fig.cap="Population structure of a set of 211 inbred lines from the tropical/subtropical groups of the maize diversity panel. ($\\textbf{A}$) Structure coefficients assuming K=4 ancestral populations. ($\\textbf{B}$) Principal Component Analysis. Individual points are colored according to the cluster with which they have the highest overlap."}
k <- 4
g1 <- slist %>%
  purrr::set_names(., change_to_k(names(.))) %>%
  purrr::map(.f = ~dplyr::mutate(., G = genos)) %>%
  purrr::keep(names(.) == k) %>%
  purrr::map(.f = ~dplyr::mutate(., Central_Cluster = Cluster1 + Cluster4)) %>%
  purrr::map(.f = ~tidyr::gather(
    ., key = "Cluster", value = "AncCoef", -G, -Central_Cluster)
    ) %>%
  dplyr::bind_rows(.id = "K") %>%
  dplyr::mutate(G = as.factor(G)) %>%
  ggplot(aes(
    x = fct_reorder(G, x = Central_Cluster), y = AncCoef, fill = Cluster)
    ) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_blank(),
    legend.position = "top"
  ) +
  guides(fill = guide_legend(ncol = 2)) +
  xlab("Genotype") +
  ylab("Ancestry Coefficient")

g2 <- cluster_df %>%
  dplyr::mutate(K = change_to_k(K)) %>%
  dplyr::filter(K == k) %>%
  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
  geom_point(size = 1) +
  guides(color = guide_legend(
    override.aes = list(size = 3),
    title = "Cluster"
    )) +
  scale_color_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside"
  )

cowplot::plot_grid(
  g1, g2, labels = c("A", "B"), nrow = 2
)
```

The set of selected individuals (Fig. \@ref(fig:pca-plot)) roughly corresponds
to clusters 1 and 4 in Fig. \@ref(fig:pca-structure-plot) and excludes
individuals in the tails of the PCA plot.

```{r pca-plot, fig.cap="Selecting individuals based on PCA"}
selected_df <- slist %>%
  purrr::set_names(., change_to_k(names(.))) %>%
  purrr::map(.f = ~dplyr::mutate(., G = genos)) %>%
  purrr::keep(names(.) == k) %>%
  .[[1]] %>%
  dplyr::left_join(y = pc_df, by = "G") %>%
  dplyr::mutate(main_cluster = dplyr::case_when(
    PC1 >= -30 & PC1 <= 10 & PC2 >= -25 & PC2 <= 40 ~ "yes",
    PC1 != 0 ~ "no"
    ))

selected_df %>%
  dplyr::mutate(main_cluster = as.factor(main_cluster))  %>%
  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
  geom_point(size = 1) +
  guides(color = guide_legend(
    override.aes = list(size = 3),
    title = "Selected"
    )) +
  scale_color_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside"
  )
```

# References