### 2017-09-01 reports/resampling_report.Rmd
send resampling pred report

---
title: Single-Step Prediction with Resampling
author: Matthias Westhues
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    self_contained: true
    toc: true
    toc_float: true
    theme: readable
    fig_caption: yes
highlight: tango
documentclass: book
bibliography: ["~/Dropbox/Paper/library.bib"]
biblio-style: apalike
link-citations: yes
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
# Data and packages -------------------------------------------------------
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  "tidyverse",
  "viridis",
  "ggthemes",
  "forcats",
  "LEA",
  "pophelper",
  "cowplot",
  "dtplyr",
  "data.table",
  "gsubfn"
  )
#devtools::install_github("mwesthues/sspredr")
pacman::p_load_gh("mwesthues/sspredr")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(
  fig.width = 8, fig.height = 6, warning = FALSE, echo = FALSE,
  message = FALSE
  )
```

# Data Preparation
We analyzed hybrid maize data from the public breeding program of the University
of Hohenheim as well as [data on inbred lines](http://www.maizego.org/Resources.html) from the lab of doctor Jianbing Yan.
The data on inbred lines were downloaded on 2017-01-31 from the [Baidu Cloud](https://pan.baidu.com/s/1eQH3hfW#list/path=%2F)
set up by the Yan lab.
The Hohenheim data were generated as part of the publication "Omics-based Hybrid
Prediction in Maize" [@Westhues2017].

## Agronomic Data
For the maize inbred lines we focused on the following traits that were
previously described in @Guo2016:

*   `Cob weight` (CW)
*   `Days to silking` (DS)
*   `Ear diameter` (ED)
*   `100-grain weight` (GW)
*   `Kernel width` (KW)
*   `Plant height` (PH)

In the case of maize hybrids we analyzed the traits

*   `Dry matter yield` (DMY)
*   `Dry matter content` (DMC)
*   `Lignin` (ADL)
*   `Fat` (FAT)
*   `Protein` (PRO)
*   `Starch` (STA)
*   `Sugar` (SUG)

## Genomic Data
The preparation of genomic marker data for hybrids and inbred lines,
respectively, involved the following five steps (for hybrid data these were
applied separately to each of the two heterotic groups):

1.   Remove loci with a callfrequency $<0.95$.

2.   Remove loci with heterozygosity $\geq 0.05$.

3.   Remove loci with minor allele frequency $<0.05$.

4.   Impute remaining missing values using BEAGLE using 25 iterations and 20
     samples per iterations.

5.   Remove copies (i.e. loci in perfect LD with a previous one) of marker
     loci.

## Transcriptomic Data
For both material types transcriptomic data were already sufficiently
pre-processed.

## Population Structure
Principal component analyses were run for the hybrid and inbred material.
The separation between the Dent and the Flint heterotic group for the hybrid
material looked as expected but for the inbred lines we found unexpected
clustering of the lines, as determined via a STRUCTURE analysis [@Pritchard2000]
and a PCA, albeit it with very low fractions of variance explained by the first
five principal components.
In total, I have examined 8 putative ancestral populations (K=2 to K=9) and,
based on combinations of bar plots and PCA plots
(Fig. \@ref(fig:pca-structure-plot)), as well as the estimated probability of
the data for each K, I chose K=4 to represent the population structure
(Fig. \@ref(fig:pca-structure-plot)).
Based on our last discussion we concluded that the pre-selection of a subset
of individuals might be necessary to mitigate the impact of population structure
on predictive ability.
The set of selected individuals entails individuals assigned to clusters 1 and 4
in Fig. \@ref(fig:pca-structure-plot) and excludes individuals in the tails of
the PCA plot.
Inbred lines that belonged to the clusters `1` and `4`, which were rather
homogeneous, were assigned to scenario `A`.
We defined a second scenario `B` which simply comprised all available inbred
lines and ran it to have predictive abilities for all genotypes in case
population structure was no issue for this material.
Finally, we defined scenario `C` as a subset of scenario `B`, comprising the
same set of genotypes as `B` but including a fixed effect for the cluster to
which each genotype's ancestry was primarily assigned.

```{r pca-structure, cache=TRUE, results='hide'}
# For the second scenario, keep only names of genotypes, which are covered by
# all data types (i.e. phenotypic, genotypic and transcriptomic).
common_genotypes <- readRDS(
  "./data/derived/maizego/unique_snp-mrna-pheno_genotypes.RDS"
)
common_genotypes <- common_genotypes %>%
  purrr::reduce(intersect)

snp <- "./data/processed/maizego/imputed_snp_mat.RDS" %>%
  readRDS() %>%
  #.[rownames(.) %in% common_genotypes, ] %>%
  sspredr::ensure_snp_quality(
    ., callfreq_check = FALSE, maf_check = TRUE, maf_threshold = 0.05,
    any_missing = FALSE, remove_duplicated = TRUE
  )

# PCA ---------------------------------------------------------------------
write.lfmm(snp, "./data/derived/maizego/snp.lfmm")

# Determine the structure of the data using genotypic data.
snp_pc <- LEA::pca(input.file = "./data/derived/maizego/snp.lfmm",
                   scale = TRUE)

pc_mat <- snp_pc$projections
rownames(pc_mat) <- rownames(snp)
colnames(pc_mat) <- paste0("PC_", seq_len(ncol(pc_mat)))

pc_df <- pc_mat %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "G") %>%
  tidyr::gather(key = PC, value = Score, -G) %>%
  tibble::as_data_frame() %>%
  dplyr::filter(PC %in% paste0("PC_", seq_len(2))) %>%
  dplyr::mutate(
    PC = gsub("PC_", replacement = "PC", x = PC)
  ) %>%
  tidyr::spread(key = PC, value = Score)

# PCA ---------------------------------------------------------------------
genos <- readr::read_tsv(
  "./data/processed/maizego/popstruc.names",
  col_names = FALSE
  ) %>%
  dplyr::select(2) %>%
  purrr::flatten_chr()

# Select K=4
# For each individual, return the cluster with the largest coefficient.
# Assign a color to each cluster.
# Plot the PCA results with colors corresponding to the cluster with the largest
# coefficient for that individual.

get_max_cluster <- function(x) {
  x %>%
    dplyr::mutate(G = genos) %>%
    tidyr::gather(key = "Cluster", value = "AncCoef", -G) %>%
    dplyr::group_by(G) %>%
    dplyr::summarize(main_cluster = which.max(AncCoef)) %>%
    dplyr::mutate(main_cluster = main_cluster %>%
      as.character() %>%
      as.factor()
    ) %>%
    dplyr::ungroup()
}

# STRUCTURE FILES ---------------------------------------------------------
slist <- list.files(
  path = "./data/processed/maizego",
  pattern = "popstruc[0-9]",
  full.names = TRUE
  ) %>%
  pophelper::readQ(
    files = .,
    indlabfromfile = TRUE,
    filetype = "structure"
  )

cluster_df <- slist %>%
  purrr::map(get_max_cluster) %>%
  purrr::map(.f = ~dplyr::right_join(., y = pc_df, by = "G")) %>%
  dplyr::bind_rows(.id = "K")

change_to_k <- function(x) {
  x %>%
    gsub(x = ., replacement = "", pattern = "_f") %>%
    gsub(x = ., replacement = "", pattern = "popstruc")
}
```

```{r pca-structure-plot, fig.height=12, fig.cap="Population structure of a set of 211 inbred lines from the tropical/subtropical groups of the maize diversity panel. ($\\textbf{A}$) Structure coefficients assuming K=4 ancestral populations. ($\\textbf{B}$) Principal Component Analysis. Individual points are colored according to the cluster with which they have the highest overlap."}
k <- 4
g1 <- slist %>%
  purrr::set_names(., change_to_k(names(.))) %>%
  purrr::map(.f = ~dplyr::mutate(., G = genos)) %>%
  purrr::keep(names(.) == k) %>%
  purrr::map(.f = ~dplyr::mutate(., Central_Cluster = Cluster1 + Cluster4)) %>%
  purrr::map(.f = ~tidyr::gather(
    ., key = "Cluster", value = "AncCoef", -G, -Central_Cluster)
    ) %>%
  dplyr::bind_rows(.id = "K") %>%
  dplyr::mutate(G = as.factor(G)) %>%
  ggplot(aes(
    x = fct_reorder(G, x = Central_Cluster), y = AncCoef, fill = Cluster)
    ) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_blank(),
    legend.position = "top"
  ) +
  guides(fill = guide_legend(ncol = 2)) +
  xlab("Genotype") +
  ylab("Ancestry Coefficient")

g2 <- cluster_df %>%
  dplyr::mutate(K = change_to_k(K)) %>%
  dplyr::filter(K == k) %>%
  ggplot(aes(x = PC1, y = PC2, color = main_cluster)) +
  geom_point(size = 1) +
  guides(color = guide_legend(
    override.aes = list(size = 3),
    title = "Cluster"
    )) +
  scale_color_brewer(palette = "Set1") +
  theme_pander(base_size = 15) +
  theme(
    legend.position = "top",
    strip.background = element_blank(),
    strip.placement = "outside"
  )

cowplot::plot_grid(
  g1, g2, labels = c("A", "B"), nrow = 2
)
```

# Predictions
## Preparation
### Previous work for this manuscript
In the first draft we wanted to evaluate the influence of the genetic
constitution of the set of genotypes that has only data on one out of two
predictors by generating core samples.
We defined core samples as a subset of genotypes that is covered by two
predictors whereas all other genotypes (*i.e.* the complement) are only covered
by a single predictor.
The first application involved assembling core sets of varying sizes by
maximizing the average genetic distance among core set members using the
Modifed Rogers (MR) distance as the criterion [@Thachuk2009].
The size of the core set was varied in increments of ten percentage points and
ranged from 10% to 90% of all inbred lines.
The issue with this approch is that we could not preclude effects of population
structure in our material on predictive abilities.
Therefore, we decided to replace the core sampling procedure by a nested random
subsampling procedure.

### Nested subsampling scheme
A first level of sampling was applied to each combination of `Material` (*i.e.*
"Hybrid" or "Inbred"), `Scenario` ("A", "B" and "C" for inbred lines; "None" for
hybrids), `Extent` ("Core" if every predictor was available for all genotypes
and "Full for the entire set of genotypes") and `Core_Fraction`, representing
the share of core-set genotypes for which data on all predictors were assumed
to be available (Fig. \@ref(fig:sampling-prediction-scheme) **a**,
Table \@ref(tab:genotype-count)).
For the variable `Core_Fraction` we artificially removed information on a
predictor for some genotypes to implore the influence of genetic space-coverage
through incomplete predictors on predictive ability.
This random subsampling procedure was repeated 20 times and stored as
`Rnd_Level2`.

```{r genotype-count, cache=TRUE}
html_table_width <- function(kable_output, width) {
  width_html <- paste0(paste0('<col width = "', width, '">'), collapse = "\n")
  sub("<table>", paste0("<table>\n", width_html), kable_output)
}

"./data/processed/predictions/predictions.RDS" %>%
  readRDS() %>%
  dplyr::mutate(Combi = substr(Interval, start = 1, stop = 3)) %>%
  dplyr::group_by(Combi) %>%
  dplyr::distinct(TST_Geno) %>%
  dplyr::count() %>%
  dplyr::rename(`Extent-Material-Scenario` = "Combi") %>%
  knitr::kable(
    x = .,
    booktabs = TRUE,
    caption = "Number of genotypes (n) for each combination of extent (F = full, C = core), material (I = inbred, H = hybrid) and scenario."
  ) %>%
  html_table_width(c(200, 100))
```

For all core-set inbred lines we applied a second, nested randomization scheme
where $n * x\%$ genotypes were declared to have both, data on the complete and
the incomplete predictor, whereas $n - n * x\%$ genotypes had only information
on the complete predictor (Fig. \@ref(fig:sampling-prediction-scheme) **b**).
Here $x\%$ denotes the fraction of all genotypes having information on both
predictors.
This second randomization was nested within each randomization stored in
`Rnd_Level2` and applied 20 times per randomization in `Rnd_Level2`.
Therefore, we ended up with 20 sets of predictions for each combination of
`"Material" * "Extent" * "Scenario"` when $x\% = 100$ and `20 ** 2 = 400`
sets of predictions for each combination of `"Material" * "Extent" *
"Scenario" * "Core_Fraction"` when $x\% \neq 100$.

```{r sampling-prediction-scheme, fig.show='hold', fig.cap="Prediction sampling scheme. ($\\textbf{a}$) Level 1 sampling scheme. During the first round of sampling, 70% of the genotypes were drawn at random (black symbols) and assigned to the training set (TRN) whereas phenotypes for the other 30% of genotypes (yellow symbols) were set to missing for this particular random sample. The random samples drawn at this level were subsequently denoted as s~1~ to s~20~. ($\\textbf{b}$) Level 2 sampling scheme. For maize inbred lines from the diversity panel, some genotypes had both genotypic and transcriptomic predictor data (black symbols). For each level of s from the first round of random sampling, a second round of random sub-sampling was applied where x% of the genotypes from the training set of s were declared to have both, genotypic and transcriptomic data (black symbols) whereas 100-x% of the genotypes were declared to have only genotypic information (gray symbols). This nested subsampling process was repeated 20 times and for nine different values of $x = (10\\%, 20\\%, 30\\%, \\dots, 90\\%)$. ($\\textbf{c}$) Leave-one-out cross-validation (LOOCV) scheme for maize diversity panel inbred lines. For each of the subsets from level 1 (and level 2) LOOCV was applied. Thereto, $n$ TST were generated (yellow symbols) where all genotypes eligible for TRN membership based on $\\textbf{a}$ were assigned to TRN (black symbols). ($\\textbf{d}$) LOOCV scheme for maize hybrids. This scheme is based solely on level 1 randomization in $\\textbf{a}$ and, for each level of s, places only T0 hybrids, which do not share a common parent with a TST hybrid, inside the TRN."}
knitr::include_graphics("compound_plot.svg", auto_pdf = FALSE)
```

## Bootstrap
Separately for each combination of `Material`, `Extent`, `Scenario`, `Trait`,
`Core_Fraction`, `Predictor`, `Rnd_Level1` and `Rnd_Level2`, a
bootstrap of 10,000 runs was applied to a data frame comprising observed and
predicted phenotypic values.
This yielded 10,000 bootstrap predictive abilities for each of the
aforementioned combinations together with a corresponding standard error.
The final predictive ability was calculated by taking the arithmetic mean
of bootstrapped predictive abilities over `Rnd_Level1` and `Rnd_Level2`.
The corresponding standard error was calculated across the average bootstrapped
standard error across `Rnd_Level1` and `Rnd_Level`.

## Visualization
```{r prediction-data, cache=FALSE, results='hide'}
boot_df <- "./data/processed/predictions/bootstrapped_predictions.RDS" %>%
  readRDS() %>%
  dplyr::group_by(Combi, Trait, Core_Fraction, Predictor) %>%
  dplyr::summarize(
    avg_r = mean(r),
    se = mean(std.error)
  ) %>%
  dplyr::ungroup()

## -- HYBRIDS -----------------------------------------------------------------
# Specify a unified predictor order.
pred_order <- c("P", "G", "T", "PG", "PT", "GT")

# Specify a unique trait order.
trait_order <- c("DMY", "DMC", "ADL", "FAT", "PRO", "STA", "SUG")

hyb_df <- boot_df %>%
  dplyr::filter(Combi %in% c("CHN", "FHN")) %>%
  dplyr::mutate(Trait = forcats::fct_recode(
    Trait,
    "DMY" = "GTM",
    "DMC" = "GTS",
    "FAT" = "FETT",
    "PRO" = "RPR",
    "SUG" = "XZ"
  )) %>%
  dplyr::mutate(Combi = forcats::fct_recode(
    Combi,
    "Core" = "CHN",
    "Full" = "FHN"
  )) %>%
  dplyr::mutate(Trait = forcats::fct_relevel(Trait, trait_order)) %>%
  dplyr::mutate(
    Predictor = gsubfn::gsubfn(
      pattern = "\\S+",
      replacement = list(
        "mrna" = "T",
        "ped" = "P",
        "snp" = "G",
        "ped_snp" = "PG",
        "ped_mrna" = "PT",
        "snp_mrna" = "GT"
      ),
      x = Predictor
    ),
    Predictor = forcats::fct_relevel(Predictor, pred_order)
  )

# Because the bars and errorbars have different widths
# we need to specify how wide the objects we are dodging are
dodge <- position_dodge(width = 0.9)

# Predefine the color scheme.
mycols <- scales::brewer_pal(type = "div", palette = "Spectral")(n = 6)
names(mycols) <- c("P", "G", "T", "PG", "PT", "GT")
limits <- aes(
  ymax = `Predictive Ability` + se,
  ymin = `Predictive Ability` - se
  )

## -- INBREDS CORE_FRACTION == 1 ----------------------------------------------
inbred_df <- boot_df %>%
  dplyr::filter(!Combi %in% c("CHN", "FHN")) %>%
  dplyr::mutate(
    Predictor = gsubfn::gsubfn(
      pattern = "\\S+",
      replacement = list(
        "snp" = "G",
        "mrna" = "T",
        "snp_mrna" = "GT"
      ),
      x = Predictor
    ),
    Trait = gsubfn::gsubfn(
      pattern = "\\S+",
      replacement = list(
        "Plantheight" = "PH",
        "Eardiameter" = "ED",
        "100grainweight" = "GW",
        "cobweight" = "CW",
        "Kernelwidth" = "KW",
        "Silkingtime" = "DS"
      ),
      x = Trait
    )
  ) %>%
  dplyr::mutate(Combi = forcats::fct_recode(
    Combi,
    "Core" = "CHN",
    "Full" = "FHN"
  ))

# Predefine the color scheme.
# Use also the levels of predictors in the hybrid data set to ensure that,
# regardless of the material, the same color is picked for the same predictor.
a_colors <- scales::brewer_pal(type = "div", palette = "Spectral")(n = 6) %>%
  set_names(c("P", "G", "T", "PG", "PT", "GT")) %>%
  .[names(.) %in% c("G", "T", "GT")]
```

```{r hybrid-plot, cache=FALSE, fig.cap="Predictive abilities for seven agronomic traits in maize hybrids based on three different predictors and combinations thereof. The bootstrapped standard error is displayed through error bars. The predictors are pedigree (P), genomic data (G) and transcriptomic data (T). Two letters indicate single step prediction where the first letter denotes the complete predictor whereas the second letter denotes the incomplete predictor. The 'Core' set of 685 hybrids entails all hybrids with records for all predictors. The 'Full' set comprises 1521 hybrids with data for P and G and the subset of 685 hybrids that additionally has information on T."}
hyb_df %>%
  dplyr::rename(`Predictive Ability` = "avg_r") %>%
  ggplot(aes(x = Trait, y = `Predictive Ability`, fill = Predictor)) +
  geom_bar(stat = "identity", position = dodge, color = "black") +
  geom_errorbar(limits, position = dodge, width = 0.25) +
  facet_grid(. ~ Combi) +
  scale_fill_manual(values = mycols) +
  ggthemes::theme_pander(base_size = 10)
```

```{r inbred-plot, cache=FALSE, fig.cap="Predictive abilities for six agronomic traits in maize inbred lines based on two different predictors and their combination. The bootstrapped standard error is displayed through error bars. The predictors are genomic data (G) and transcriptomic data (T). Two letters indicate single step prediction where the first letter denotes the complete predictor and the second letter denotes the incomplete predictor. Different combinations of extent (C = 'Core' and F = 'Full') with prediction scenario (A, B, or C) are plotted separately along the y-axis where the fist letter indicates the extent of the material and the third letter denotes the scenario. The 'Core' set of 149 inbred lines entails all genotypes with records for both predictors. The 'Full' set comprises 211 genotypes with complete data for G and the subset of 149 inbred lines that additionally has information on T. Note that scenario A only entails the 110 inbred lines that belong to clusters 1 and 4 in the PCA (Fig. 3.5)."}
inbred_df %>%
  dplyr::filter(Core_Fraction == "1") %>%
  dplyr::rename(`Predictive Ability` = "avg_r") %>%
  ggplot(aes(x = Trait, y = `Predictive Ability`, fill = Predictor)) +
  geom_bar(stat = "identity", position = dodge) +
  geom_errorbar(limits, position = dodge, width = 0.25) +
  facet_grid(Combi ~ .) +
  scale_fill_manual(values = a_colors) +
  theme_pander(base_size = 10)
```

```{r core-inbred-plot, cache=FALSE, fig.cap="Influence of the size of the subset of genotypes with both, genomic and transcriptomic data on predictive abilities. Results are shown for six agronomic traits in the core set of 149 maize inbred lines ('CIB' and 'CIC') and the core set of 110 genotypes ('CIA'), respectively. The fraction of genotypes with information on both predictors is denoted as 'Core Fraction' and displayed on the x-axis."}
inbred_df %>%
  dplyr::filter(Predictor == "GT", Combi %in% c("CIA", "CIB", "CIC")) %>%
  dplyr::mutate_at(vars(Predictor), funs(as.factor)) %>%
  dplyr::rename(`Core Fraction` = "Core_Fraction") %>%
  dplyr::rename(`Predictive Ability` = "avg_r") %>%
  ggplot(aes(
    x = `Core Fraction`,
    y = `Predictive Ability`,
    color = Trait,
    group = Trait
    )
  ) +
  geom_line(stat = "identity", position = position_dodge(width = 0.3)) +
  facet_grid(~ Combi) +
  ggthemes::scale_color_tableau() +
  ggthemes::theme_pander(base_size = 10) +
  theme(legend.position = "right") +
  guides(color = guide_legend(override.aes = list(size = 3)))
```

# References