### 2017-10-16 paper/manuscript.tex
paper, edits from meeting on 2017_10_12

\documentclass[12pt,titlepage]{article}

%% THE USEPACKAGES NECESSARY FOR THIS EXAMPLE
%% NOTE THAT genetics_manu_style MUST BE CALLED AFTER mychicago
\usepackage{graphicx}
\usepackage{endfloat}
\usepackage{amsfonts}
\usepackage{mychicago}
\usepackage{subfigure}
\usepackage{genetics_manu_style}
\usepackage{authblk} % improved citations
\usepackage{setspace} %double spacing
\usepackage{lineno} % line numbers
\usepackage{booktabs} % book-quality tables
\usepackage{multicol} % list corresponding authors in two separate columns
\usepackage{hyperref} % web links
\usepackage{tensor} % tensor matrices

%% Probably unnecessary packages.
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript

%% THE MANUSCRIPT TITLE
\title{Efficient Phenotype Prediction Using Incomplete Omics Data}

% Allow for multiple authors to share the same institution.
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

\author{
  Matthias Westhues\thanks{Institute of Plant Breeding, Seed Science and Population Genetics, University of Hohenheim, 70599 Stuttgart, Germany},
  Claas Heuer\thanks{Institute of Animal Breeding and Husbandry, Christian-Albrechts-University Kiel, 24098 Kiel, Germany},
  Georg Thaller\samethanks[2],
  Rohan Fernando\thanks{Department of Animal Science, Iowa State University, 50011 Ames, Iowa, U.S.A.},
  Albrecht E. Melchinger\samethanks[1]
}

\renewcommand{\CorrespondingAddress}{%
    University of Hohenheim \\
    Institute of Plant Breeding, Seed Science and Population Genetics \\ 
    Fruwirthstr. 21 \\ 
    Stuttgart 70599, GERMANY \\
    Tel.: (+49) 0711 459--22334 \\
    Fax: (+49) 0711 459--22343 \\ 
    \texttt{melchinger@uni-hohenheim.de} \vfill}

\renewcommand{\RunningHead}{Single Step Prediction}
\renewcommand{\CorrespondingAuthor}{Prof. Dr. A. E. Melchinger}
\renewcommand{\KeyWords}{single step, omics, prediction, maize}

%% SOME COMMANDS FOR THE CONTENT OF THIS FILE.  NOT NECESSARY FOR
%% GENETIC_MANU_STYLE
\newcommand{\bp}{\mathbf{p}}
\newcommand{\LLL}{\mathcal{L}}

%% BEGIN DOC
\begin{document}

\maketitle
\doublespacing
\linenumbers

\section{Key message}
Borrowing information from a highly performant `omics' predictor, sampled on a
subset of inbred lines, can efficiently improve the prediction of complex
traits.

\begin{abstract}
Predicting genotypic values with great precision is pivotal for effective
selection of candidates in animal and plant breeding.
Novel `omics' predictors have shown to improve upon established genome-based
predictions of some complex traits but yet require laborious and expensive
assays.
In animal breeding, single-step prediction was introduced to efficiently use
pedigree information, collected on a large number of animals, and genomic
information, collected on a smaller subset of animals, simultaneously for breeding
value estimation without bias.
Single-step prediction has not yet been attempted with predictors calculated from
quantitative features displaying repeatabilities smaller than one.
Using two maize datasets of inbred lines and hybrids we show that transcriptomic
data can be imputed successfully, boosting forecasts when their predictive
ability exceeds that of pedigree or genomic data.
Our results suggest that covering only a subset of inbred lines with `omics'
predictors and imputing all others using pedigree or genomic data could enable
breeders to improve trait predictions while keeping costs under control.
Employing `omics' predictors could particularly improve candidate selection in
hybrid breeding where the success of forecasts is a strongly convex function of
predictive ability.
\end{abstract}

% Cover Letter
\section{Cover Letter}
Dear Managing editors of TAG,

We wish to submit the enclosed manuscript 
``Efficient Phenotype Prediction Using Incomplete Omics Data'' by Westhues et 
al.\@ for publication in TAG.

Recent studies in maize inbred lines~\cite{Guo2016}, maize hybrids
\cite{Zenke-Philippi2017,Westhues2017} and rice hybrids~\cite{Dan2016,Xu2016}
have suggested that `omics' predictors, such as the metabolome and the
transcriptome, can improve the prediction of traits with complex genetic 
architectures over that achieved with pedigree or genomic data.
Particularly hybrid breeding, where the success of forecasts increases
exponentionally with predictive abilities, could benefit from the use of
predictors that yield more precise estimates of candidate performance.
Currently, the adoption of these novel predictors by breeders is impeded
by their higher costs compared to pedigree and genomic data.
Here, we present single-step prediction of complex traits as a framework for
combining pedigree and/or genomic data with other `omics' predictors collected
on a subset of all available inbred lines to boost predictive abilities while
controlling costs.

Using two extensive phenotypic data sets from a public hybrid silage maize
breeding program and a maize diversity panel published in TAG~\cite{Westhues2017,
Guo2016}, we a) investigate the success of imputing a quantitative predictor
with feature repeatabilities less than one and b) the influence of the fraction
of inbred lines covered by two predictors on predictive abilities in prediction
based on incomplete `omics' data.

In our current study we obtained the following results:

\begin{enumerate}
\item In both data sets, single-step prediction improved predictive abilities
	over those of the predictor covering all inbred lines when the
	discrepancy between the	two predictors in terms of predictive ability
	were moderate to large in favor of the incomplete predictor.
\item When 20\%-30\% ($n \approx 25$) of the inbred lines were covered by both
	predictors, single step-based predictive abilities converged to those
	obtained with the incomplete predictor.
\end{enumerate}

In conclusion our results demonstrate that single-step prediction, using
an established predictor such as pedigree information, can benefit from sampling
a highly performant predictor, such as transcriptomic data, on a subset of all
inbred lines.
The framework presented herein should help promote the use of attractive
`omics' predictors by balancing their advantages with their currently higher
costs.

% Introduction
\section{Introduction}
% Rationale for predictions.
Genomic prediction, pioneered by \citeNP{Meuwissen2001}, has revolutionized
animal and plant breeding by a more efficient selection of promising candidates 
without phenotypic records~\cite{DeLosCampos2013,Garcia-Ruiz2016}.
Its use is mainly motivated by the high costs and oftentimes time-consuming
phenotyping.
Costs can be reduced when all genotypes, equivalent to inbred lines in plant
breeding and animals in animal breeding, are covered by a predictor while
limiting phenotyping to a few genotypes with high precision \cite{Kadam2016}.
The performance of all other genotypes is then forecast using model parameters
estimated in the subset of individuals with full information.
In animal breeding, genomic selection has increased the annual selection gain 
by $\approx$ 50--100\% for yield traits and by 300--400\% for traits with low 
heritability~\cite{Garcia-Ruiz2016}.

% Incomplete predictors.
Breeding programs --- particularly in animal breeding --- are characterized by
the availability of one predictor type, which is available for all recorded
genotypes with phenotypic information while other predictors are incomplete in
that they cover only a subset of all genotypes~\cite{Fragomeni2015}.
Usually, the complete predictor is pedigree information while genomic data are
the most prevalent incomplete predictor.
To utilize all available phenotypic information for model training, methods were 
developed that allow the combined use of genomic and pedigree information 
\cite{Hayes2009a,VanRaden2009}.

% Two-step prediction.
Initially, a two-step procedure was the most intuitive choice for accomplishing
this goal~\cite{VanRaden2009}.
The first step consists of conventionally estimating breeding values based on
pedigree relationships among the candidates and phenotypes of their relatives.
Individuals with higly reliable breeding values out of the first step,
(\textit{i.e.} those having a high number of offspring) enter the second step in
which conventional breeding values are regressed onto SNP genotypes.
The estimated model coefficients can then be used to predict breeding values
for solely genotyped and potentially very young individuals.
However, as the breeding values from step one enter the genomic prediction
model as phenotypes, one has to be careful about the residual error
distribution of such a model.
For instance, the range of the prediction errors will no longer be identically
distributed and thus, they should be weighted differently or estimated 
individually~\cite{Aguilar2010}.
One popular approach to tackle this problem is deregressing the breeding
values and weighting the residuals according to the prediction error variances
\cite{Garrick2009}.

% Workings and advantages of the single step procedure.
\citeNP{Legarra2009} and \citeNP{Christensen2010} independently from each other
developed a single-step BLUP method that blends the numerator relationship matrix
$\mathbf{A}$ with the genomic relationship matrix $\mathbf{G}$ in a mutual matrix
$\mathbf{H}$ to use all predictor information simultaneously.
Issues with this H-BLUP approach were the compatability of $\mathbf{A}$ with
$\mathbf{G}$~\cite{Christensen2012} and their weighting, which is not
trait-independent~\cite{Vitezica2011,Ashraf2016}.
\citeN{Fernando2014} derived the equivalent single-step marker-effect-model to 
the single-step breeding value model of \citeNP{Legarra2009} and 
\citeNP{Christensen2010}, which does not require weights for $\mathbf{A}$ and 
$\mathbf{G}$ and allows implicit modeling of the imputation error structure.

% -   Hybrids
In animal breeding, single-step prediction, particularly H-BLUP, has already 
become a routine procedure~\cite{Legarra2014} for predicting breeding values
but, to the best of our knowledge, has only been considered in plants in a study
on wheat~\cite{Ashraf2016} and white spruce lines, respectively
\cite{Ratcliffe2017}.
Moreover, while an extension of the single-step framework to hybrid genotypes
was developed for pigs~\cite{Xiang2016}, it is yet lacking for the prediction
of single-cross hybrids between purely inbred lines.
Technological advances in hybrid breeding of several crop species, such as the
rapid generation of completely homozygous lines~\cite{Wedzony2009}, now enable
plant breeders to produce about $n = 1,000$ individuals per generation and sex.
Assuming that about 90\% of these lines were never before tested in any hybrid
combination about $n^{2} * 0.9^{2} = 810,000$ putative hybrids per generation
would not have a single parent with phenotypically tested progeny
\cite{Westhues2017}.
Predicting the performance of these hybrid candidates with low relatedness to 
tested hybrids using only information from their parents is currently a hot topic
in plant breeding~\cite{Technow2014,Kadam2016,Westhues2017}.

% -   quantitative predictores
Given that the availability of pedigree and genomic data is the most prevalent
situation in animal breeding, previous studies have focused exclusively on the
application of single-step prediction to these predictors.
Recent investigations into the utility of `omics' data for boosting predictive
abilities in plants have suggested that metabolic and transcriptomic data can
improve upon predictive abilities based on pedigree or genomic data
\cite{Guo2016,Xu2016,Dan2016,Zenke-Philippi2017,Westhues2017}.
Yet, these two `omics' technologies are not fully mature and therefore incur
higher costs to breeders than pedigree and genomic data, putting their utility
into question despite their proven usefulness.

% - Objectives
Hence, our objectives were to (i) extend the existing theoretical framework for
single-step prediction to the imputation of quantitative `omics' features, here
exemplified by transcriptomic data, (ii) evaluate the impact of the proportion of
imputed genotypes in the training set on the predictive ability and (iii) transfer
the single-step framework for breeding value prediction to the prediction of
hybrids.

\section{Materials and Methods}
\Genetics2level{Experiment 1}
To address objectives (i) and (ii) a published data set of maize inbred lines
\cite{Yang2014} was used.
It originally comprised 513 maize lines representing the global maize
diversity  and was reduced to the set of tropical and subtropical lines
($n = 211$), subsequently referred to as Exp1, which is the largest of the four
subgroups classified by~\citeN{Guo2016}.
All inbred lines were evaluated in five different environments in China
described in detail by~\citeNP{Yang2014}.
Best linear unbiased predictors (BLUPs) were calculated by~\citeN{Yang2014} for
all 211 inbred lines and for 17 traits of which six are presented in this study.
The selection of these traits was done based on our prediction results so that
two traits were predicted with greater precision by transcriptomic data and four
further traits were predicted with greater precision by genomic data.
After filtering for minor allele frequency ($\geq 5$\%), heterozygosity rate
($\leq 5$\%) and call frequency ($\geq 95$\%), missing genotypes were imputed
using the Beagle software~\cite{Browning2009} resulting in 37,760 SNPs for these
211 genotypes.
A subset of 28,850 annotated genes was available for 149 out of the 211
genotypes~\cite{Fu2013} and provided the basis for further analyses.

\subsection{Population structure}
Population structure in Exp1 was examined through a STRUCTURE
\cite{Pritchard2000} analysis with the number of putative ancestral
populations $K$ ranging from 2 to 9.
Based on the estimated likelihood of the data for each $K$, we chose $K=4$
and carried out a principal component analysis (PCA) on the scaled and centered
genomic feature matrix using the \emph{pca} function from the \emph{R}-package
\emph{LEA}~\cite{Frichot2015}.
Least-squares estimates of ancestry proportions were estimated based on sparse 
nonnegative matrix factorization algorithms~\cite{Frichot2014} using the 
\emph{snmf} function in \emph{LEA}.
Genotypes in the PCA plot (Fig.~\ref{fig:inbred-pca}) were colorized based on their
primary assignment to any of the four putative ancestral populations.
To mitigate the possibility of inflated predictive abilities, which can arise
from large differences in the average performance of the different
subpopulations~\cite{Windhausen2012}, we reduced the panel of inbred lines to
164 genotypes, which were assigned to the largely overlapping clusters 1 and 4 in
Fig.~\ref{fig:inbred-pca}.
In summary, genomic and phenotypic information were available for 164 inbred
lines, which we denote as the `Full' set, whereas transcriptomic data were
available for a subset of 110 inbred lines, which we denote as `Reduced'.

\begin{figure}[H]
  \centering
  \includegraphics{./tables_figures/inbred_pca.pdf}
  \caption{
  Principal component analysis (PCA) of 211 genotypes from the set
  `tropical/subtropical' of the maize diversity panel inbred lines.
  The four clusters are based on $K=4$ putative ancestral populations from a
  STRUCTURE analysis of the corresponding genomic feature matrix.
  Genotypes are colored based on their primary assignment to any of these four
  clusters.
  The explained variance for each of the two PCs is given in the corresponding
  axis labels.
}
\label{fig:inbred-pca}
\end{figure}

\Genetics2level{Experiment 2}
To address objectives (i) and (iii), a maize hybrid data set, subsequently
denoted as Exp2, was used.
The material comprised 1,521 hybrids produced in 16 factorial mating designs
between 142 Dent and 103 Flint parent lines described in detail by
\citeN{Westhues2017}.
Best linear unbiased estimates (BLUEs) were computed across all factorials and 
three or more agro-ecologically diverse environments across Germany for seven
agronomically important traits in silage maize by accounting for year, location,
field replication, block and genotype effects as well as their interactions.
All 245 parent lines have pedigree records back to the generation of their 
grandparents or deeper~\cite{Westhues2017} and were genotyped using the Illumina 
BeadChip MaizeSNP50~\cite{Ganal2011}.
The same SNP quality checks as for the material in Exp1 were applied here,
yielding 7,013 polymorphic marker loci for the Dent and 6,212 for the Flint
lines, respectively.
Gene expression data for seedlings seven days after sowing were obtained for a 
subset of 60 Dent and 43 Flint lines, that are parents of 685 hybrid progeny.
These data were obtained through two-color hybridizations using a custom 2K 
microarray (GPL22267)~\cite{Westhues2017}.
After applying established normalization procedures~\cite{Smyth2003,Ritchie2007},
BLUEs for gene expression were computed for 1,323 transcripts
\cite{Westhues2017} using the R-package \emph{limma}~\cite{Ritchie2015a}.

\Genetics2level{Prediction}
\subsection{Kernels}
Depending on the data set, up to three predictors were available for agronomic
trait predictions, namely pedigree data (P), genomic data (G) and
transcriptomic data (T).
The corresponding feature matrix of the inbred lines ($\mathbf{M}$) has
dimensions $n \times p$ where $n$ pertains to the number of genotypes and $p$
pertains to the number of features; that is transcripts for T and SNPs for G.
For the sake of simplicity the derivations are detailed first for predicting
breeding values of inbred lines.
Later, we briefly describe the extension to models required for predicting the
genetic value of hybrids.
For the complete predictor we generated kernels by centering and standardizing
all features in $\mathbf{M}$ to unit variance~\cite{VanRaden2008}:

\begin{equation} \label{eq:GenomicRelationship}
  \mathbf{K} = \frac{1}{p} \mathbf{M} \mathbf{M}^{\top}.
\end{equation}

In the case of pedigree data (P), the numerator relationship matrix was used
directly for $\mathbf{K}$.

\subsection{Genetic model}
The general model for breeding values of the lines was:

\begin{equation} \label{eq:KBLUPModel}
  \mathbf{y} = \mu + \mathbf{Z} \mathbf{g} + \mathbf{e},
\end{equation}

where $\mathbf{y}$ is the vector of observed inbred line performance $\mu$ is
the fixed model intercept, and $\mathbf{e}$ is the residual error.
For the inbred lines we fit random genotype effects $\mathbf{g}$, which are
associated with the vector of phenotypic performance $\mathbf{y}$ through a
corresponding design matrix $\mathbf{Z}$.
The random effects ($\mathbf{g}$) have expectation zero and variance equal to
$\mathbf{K} \sigma^{2}_{g}$ and $\mathbf{I} \sigma^2_{e}$ for the residual
error, respectively.
All prediction models were implemented using the R-packages \textit{BGLR}
\cite{Perez2014} and \textit{sspredr}
\url{https://github.com/mwesthues/sspredr}.

\subsection{Single-step prediction}
If the breeding values are fully captured by the features in $\mathbf{M}$,
Eq.(~\ref{eq:KBLUPModel}) can be expressed as~\cite{Fernando2014}

\begin{align} \label{eq:KBLUPModelSimplified}
	&= \mu + \mathbf{Z} \mathbf{M} \boldsymbol{\alpha} + \boldsymbol{e},
\end{align}

with $\boldsymbol{\alpha}$ being the vector of random, partial-regression
coefficients of the feature covariates, because the ridge regression model is
equivalent to the BLUP model \cite{Ruppert2003}.

Hence, the breeding values from the prediction model are given by:
\begin{equation} \label{eq:mrnaebv}
	\mathbf{\hat{g}} = \mathbf{M}\boldsymbol{\hat{\alpha}},
\end{equation}

where $\boldsymbol{\hat{\alpha}}$ is the solution to the ridge regression model.

Consider now the situation in which one predictor is complete in the sense that
the whole population is covered by features for that predictor but only a subset
of that population has information for a second predictor.
We can take a similar route as in~\citeNP{Fernando2014} and impute covariates 
of the incomplete predictor, denoted as `predictor' and indicated through the
asterisk symbol `*', by using covariates of the complete predictor, denoted as
`imputor'.
Let the subscript $1$ denote individuals covered only by the predictor whereas
individuals covered by both, the imputor and the predictor, are indicated by the subscript $2$.

Let us further denote the matrix of centered covariates of the predictor as
\begin{equation} \label{eq:incomplete-covariates}
\tensor[]{\mathbf{W}}{^{*}} =
\begin{bmatrix}
  \mathbf{Z}_{1} \; \tensor[]{\mathbf{\hat{W}}}{^{*}_{1}} \\
  \tensor[]{\mathbf{W}}{^{*}_{2}}
 \end{bmatrix},
\end{equation}
where observed feature covariates are stored in $\tensor[]{\mathbf{W}}{^{*}_{2}}$
while the feature covariates for genotypes not covered by the predictor can be
imputed with the aid of the covariance matrix of the imputor $\mathbf{K}$ and
the feature matrix of the predictor $\tensor[]{\mathbf{W}}{^{*}_{2}}$ by
calculating
($\tensor[]{\mathbf{\hat{W}}}{^{*}_{1}} = \tensor[]{\mathbf{K}}{_{12}} \;
\tensor[]{\mathbf{K}}{_{22}^{-1}} \; \tensor[]{\mathbf{W}}{^{*}_{2}}$).
In this case we need to partition the vector of breeding values $\mathbf{g}$
into a component $\mathbf{g}_{1}$ (\textit{i.e.,} breeding values of genotypes
covered only by the imputor) and a component
$\mathbf{g}_{2} = \tensor[]{\mathbf{W}}{^{*}_{2}} \; 
\tensor[]{\boldsymbol{\alpha}}{^{*}}$ (\textit{i.e.,} genetic
values of genotypes covered by both predictors, where
$\tensor[]{\boldsymbol{\alpha}}{^{*}}$ pertains to the random feature effects of
the predictor).
Features are distributed according to a multivariate normal distribution:
\begin{align} \label{eq:mrna1}
	\mathbf{g_1} &= E(\mathbf{g}_1|\mathbf{g}_2) +
	\boldsymbol{\epsilon} \\
	&= \tensor[]{\mathbf{K}}{_{12}} \;
       	\tensor[]{\mathbf{K}}{_{22}^{-1}} \;
       	\tensor[]{\mathbf{W}}{^{*}_{2}} \;
	\tensor[]{\boldsymbol{\hat{\alpha}}}{^{*}} +
       	(\mathbf{g_1} - \tensor[]{\mathbf{K}}{_{12}} \;
	\tensor[]{\mathbf{K}}{_{22}^{-1}} \;
       	\tensor[]{\mathbf{W}}{^{*}_{2}} \;
	\tensor[]{\boldsymbol{\hat{\alpha}}}{^{*}}) \\
	&= \mathbf{\hat{g}}_1 + \boldsymbol{\epsilon},
\end{align}

where the first summand pertains to $\mathbf{\hat{g}}_{1}$, the second summand
pertains to $\boldsymbol{\epsilon}$ and where the partitions of the matrix
\begin{equation} \label{eq:imputor}
\tensor[]{\mathbf{K}}{} =
\begin{bmatrix}
  \mathbf{K}_{11} & \mathbf{K}_{12} \\
  \mathbf{K}_{21} & \tensor[]{\mathbf{K}_{22}}{^{*}}
 \end{bmatrix},
\end{equation}

correspond to $\mathbf{g}_{1}$ and $\mathbf{g}_{2}$, respectively
\cite{Fernando2014} and $\mathbf{\hat{g}}_{1}$ pertains to breeding values
estimated by imputing the predictor.

The covariance matrix of $\boldsymbol{\epsilon}$ is
$\tensor[]{\mathbf{K}}{_{11}} -
\tensor[]{\mathbf{K}}{_{12}} \;
\tensor[]{\mathbf{K}}{_{22}^{-1}} \;
\tensor[]{\mathbf{K}}{_{21}} =
(\tensor[]{\mathbf{K}}{^{11}})^{-1}$
\cite{Legarra2009}

and the covariance of $\mathbf{\hat{g}}_{1}$ is
$\tensor[]{\mathbf{K}}{_{12}} \;
\tensor[]{\mathbf{K}}{_{22}^{-1}} \;
\tensor[]{\mathbf{W}}{^{*}_{2}} \;
\tensor[]{\mathbf{W}}{^{*}_{2}^{\top}} \;
\tensor[]{\mathbf{K}}{_{22}^{-1}} \;
\tensor[]{\mathbf{K}}{_{21}} \;
\tensor[]{\sigma^{2}_{\alpha}}{^{*}}$
\cite{Fernando2014}.

The structure of the residual imputation error is known, and therefore, can be
modeled.

The general model for the single-step procedure can then be written as

\begin{equation} \label{eq:single-step-model}
\mathbf{y} =
\mathbf{Xb} +
\tensor[]{\mathbf{W}}{^{*}} \;
\tensor[]{\boldsymbol{\alpha}}{^{*}} +
\mathbf{U} \boldsymbol{\epsilon} +
\mathbf{e},
\end{equation}

with

\begin{equation} \label{eq:single-step-submatrices}
\mathbf{X} =
\begin{bmatrix}
  \mathbf{X}_1 \\
  \mathbf{X}_2
 \end{bmatrix},
 \mathbf{U} =
\begin{bmatrix}
  \mathbf{Z}_1 \\
  0
 \end{bmatrix}.
\end{equation}

According to the notation in \citeNP{Fernando2014} the phenotypes can be
untangled as:

\begin{align} \label{eq:entangled-augmented-single-step-model}
\begin{bmatrix}
  \mathbf{y}_1 \\
  \mathbf{y}_2
 \end{bmatrix}
& =
 \begin{bmatrix}
  \mathbf{X}_1 \\
  \mathbf{X}_2
 \end{bmatrix}
 \boldsymbol{\beta} +
 \begin{bmatrix}
  \mathbf{Z}_1 & \boldsymbol{0} \\
  \boldsymbol{0} & \mathbf{Z}_2
 \end{bmatrix}
\begin{bmatrix}
  \mathbf{g}_1 \\
  \mathbf{g}_2
 \end{bmatrix}
  + \mathbf{e} \\
    & =
 \begin{bmatrix}
 \mathbf{X}_1 \\
 \mathbf{X}_2
 \end{bmatrix}
 \boldsymbol{\beta} +
 \begin{bmatrix}
  \mathbf{Z}_1 & \boldsymbol{0} \\
  \boldsymbol{0} & \mathbf{Z}_2
 \end{bmatrix}
\begin{bmatrix}
  \tensor[]{\mathbf{K}}{_{12}} \; \tensor[]{\mathbf{K}}{_{22}^{-1}} \;
  \tensor[]{\mathbf{W}}{^{*}_{2}} \;
  \tensor[]{\boldsymbol{\alpha}}{^{*}} + \boldsymbol{\epsilon}  \\
  \tensor[]{\mathbf{W}}{^{*}_{2}} \; \tensor[]{\boldsymbol{\alpha}}{^{*}} \\
 \end{bmatrix}
  + \mathbf{e} \\
    & =
 \begin{bmatrix}
  \mathbf{X}_1 \\
  \mathbf{X}_2
 \end{bmatrix}
 \boldsymbol{\beta} +
 \begin{bmatrix}
  \mathbf{Z}_1 & \boldsymbol{0} \\
  \boldsymbol{0} & \mathbf{Z}_2
 \end{bmatrix}
\begin{bmatrix}
  \tensor[]{\mathbf{\hat{W}}}{^{*}_{1}} \;
  \tensor[]{\boldsymbol{\alpha}}{^{*}} + \boldsymbol{\epsilon} \\
  \tensor[]{\mathbf{W}}{^{*}_{2}} \; \tensor[]{\boldsymbol{\alpha}}{^{*}} \\
 \end{bmatrix}
  + \mathbf{e} \\
\end{align}

The model above indicates that single-step prediction models ultimately use
only the matrix of known and imputed feature covariates $\mathbf{W}^{*}$ for the
estimation of breeding values while accounting for the imputation error
$\boldsymbol{\epsilon}$.

Our final model for the predictor is now:

\begin{align} \label{eq:final-model}
\mathbf{y} &= \mathbf{Xb} +
\mathbf{Z} \;
\tensor[]{\mathbf{W}}{^{*}} \;
\tensor[]{\boldsymbol{\alpha}}{^{*}} +
\mathbf{Z}\mathbf{U} \boldsymbol{\epsilon} +
\mathbf{e}.
\end{align}

Breeding values can be estimated as:

\begin{equation} \label{eq:breeding-values}
\hat{\mathbf{g}} =
 \begin{bmatrix}
  \tensor[]{\hat{\mathbf{W}}}{^{*}_{1}} \\
  \tensor[]{\mathbf{W}}{^{*}_{2}}
 \end{bmatrix}
 \tensor[]{\hat{\boldsymbol{\alpha}}}{^{*}}
 +
 \begin{bmatrix}
  \mathbf{Z}_1 \\
  0
 \end{bmatrix}
 \hat{\boldsymbol{\epsilon}}
\end{equation}

The predicted agronomic performance of the inbred lines is

\begin{equation} \label{eq:inbred-predicted-performance}
\hat{\mathbf{y}} = \mathbf{Z}\mathbf{\hat{g}} 
\end{equation}

The previous models can simply be extended to hybrids by defining separate
feature matrices for each predictor and each heterotic group; here
$\mathbf{M}_{D}$ for Dent and $\mathbf{M}_{F}$ for Flint parent lines in Exp2.
The general model for the calculation of genetic values of hybrids in Exp2 is
then

\begin{equation} \label{eq:hybrid-genetic-value-model}
  \mathbf{y} = \mu + \mathbf{Z}_{D} \mathbf{g}_{D} +
  \mathbf{Z}_{F} \mathbf{g}_{F} +
  \mathbf{e},
\end{equation}

corresponding to Eq.~\ref{eq:KBLUPModel}.

\subsection{Predictive ability and model validation}
For the validation of our predictions, we employed leave-one-out
cross-validation (LOOCV) routines.
In the case of the diversity panel maize inbred lines, LOOCV was performed by 
using a single genotype as a hold-out sample, which will be predicted by using
all other $164 - 1$ inbred lines for model training.
This process is repeated until all $164$ inbred lines have been used once for
testing and $164 - 1$ times for model training.
For the single-step prediction models, imputation was performed on the
transcriptomic data.

For the hybrid data, LOOCV was carried out as follows:
Let $D$ and $F$ denote the set of parental inbred lines from the Dent and the 
Flint group, respectively.
The set $H$ is the subset of 1,521 phenotyped hybrids within the set of all
possible hybrids $D \times F$, which would comprise $14,626 = 142 * 103$
hybrids.
Let $i \times j$ denote the hybrid to be predicted.
We can then define the parents of the training set hybrids as
$D_{i} = D \setminus \{i\}$ and $F_{j} = F \setminus \{j\}$, respectively.
The training set is then defined as 
$H_{TRN}(i \times j) = H \cap (D_{i} \times F_{j})$.
For the single-step prediction models, imputation was performed on either the
genomic or the transcriptomic data.

We judged the performance of each model by looking at its predictive ability,
which was calculated as the correlation $\rho(\mathbf{y}, \mathbf{\hat{y}})$,
where $\mathbf{\hat{y}}$ is the vector of predicted values for all genotypes.

Standard errors for the predictive abilities were calculated by bootstrapping
the predicted and observed values 1,000 times using the \emph{R} package 
\emph{boot}~\cite{Canty2017}.

\Genetics2level{Core sampling}
In order to evaluate how prediction accuracy is influenced by the fraction of
genotypes not covered by the incomplete predictor on predictive ability we
generated core sets of nine different sizes from Exp1
(Fig.~\ref{fig:compound-plot}~a).
The base of each of these core sets was the set of 110 genotypes that were
covered by both, genomic and transcriptomic data.
A core set was created by artificially declaring a pre-specified fraction 
$x = {10\%, 20\%, \dots, 90\%}$ of genotypes as lacking of transcriptomic 
information.
For each value of $x$ we created 100 core sets $s \in {1, 2, 3, \dots, 100}$ by
repeatedly sampling genotypes, whose trancsriptomic information was declared
as missing, at random (Fig.~\ref{fig:compound-plot}~b).
The artificially removed transriptomic information on these genotypes was then
imputed.
For the each core set, comprising 110 genotypes, predictive abilities were again
computed as described above, using LOOCV and bootstrapping.

\begin{figure}[H]
  \centering
  \includegraphics[width=1\textwidth]{./tables_figures/compound_plot.pdf}
  \caption{
  (\textbf{a}) Scheme for the amount of genotypes covered by predictors.
  The two extended bars represent the entire set of genotypes in either
  Exp1 or Exp2.
  All genotypes are covered by at least one predictor, which is called the
  `complete' predictor and depicted as a white bar.
  The second bar shows that only a fraction of genotypes is covered by both,
  the `complete' and the `incomplete' predictor (dark gray) and that the
  difference between the predictors will be imputed (light gray fraction).
  The four bars to the right outline the principle of generating core sets for
  Exp1.
  For nine different values of $x$, which represents the fraction of genotypes
  covered by the `incomplete' predictor, core sets are build.
  (\textbf{b}) Subsampling within core sets.
  Each of the white symbols represents a different genotype.
  For each value of $x$, 100 subsets $s$ are generated where $x\%$ of genotypes
  are randomly declared to be covered by both, the `complete' and the
  `incomplete' predictor (dark gray) whereas the remaining $100\% - x\%$ of
  genotypes are declared to be covered only by the `complete' predictor (light
  gray).
  }
\label{fig:compound-plot}
\end{figure}

\Genetics2level{Data availability}
The code for the analyses can be accessed from 
\url{https:://github.com/mwesthues/single\_step} and
\url{https://github.com/mwesthues/sspredr}.
All data from Exp1 have been published before and are publicly available.
Data for Exp2 can be made available upon request.

\section*{Results}
\Genetics2level{Experiment 1}
At the level of the 110 genotypes for which both, genomic (G) and 
transcriptomic (T) information were available, the use of T yielded higher
predictive abilities than the use of G for two out of six traits
(Fig.~\ref{fig:InbredResults}A).
In terms of predictive ability, the absolute advantage of T over G was as large 
as 0.15, observed for the trait ear diameter\@.
The predictive ability for predictions based on G were always higher for the
full set of inbred lines ($n = 164$) than for the reduced set of inbred lines
($n = 110$).
The combination of G and T in a single step (\textit{i.e.}, $T^{*}|G$, where
$G$ is the imputor and $T^{*}$ is the predictor, for the prediction of all 164
lines, performed at least as well as G alone when T was superior over G for the
reduced set of 110 lines.
In the case of kernel width and ear diameter, the use of $T^{*}|G$ yielded a slight 
improvement in predictive ability over G.
For the other four traits, where T was a worse predictor for the reduced set of
110 lines than G, predictive abilities of $T^{*}|G$ for the full set of 164
lines were also worse than those based on G alone.

\begin{figure}[H]
  \centering
  \includegraphics{./tables_figures/pred_ability_inbred.pdf}
  \caption{
  Predictive abilities and average, bootstrapped standard errors for
  tropical/subtropical inbred lines from the maize diversity panel and six
  agronomic traits.
  As predictors, genomic (G) data, transcriptomic (T) data and their 
  combination ($T^{*}|G$) were used.
  The `Reduced' panel includes a subset of 110 lines covered by genomic and
  transcriptomic information whereas the `Full' panel comprises all 164
  genotypes.
  In the `Full' set of genotypes, transcriptomic records were imputed for 54
  lines.
  }
\label{fig:InbredResults}
\end{figure}

\Genetics2level{Experiment 2}
In the `Reduced' set of Exp2, transcriptomic data were the best single predictor
for the traits lignin, protein and particulaly for dry matter yield where the
predictive ability obtained with T was 0.08 points higher than for G as the
second best predictor for this trait (Fig.~\ref{fig:HybridResults}).
In the reduced set, P had the highest predictive ability for starch and dry
matter content\@.
For fat and sugar, G was the superior predictor inside the reduced set.
In the full set of 1,521 hybrids, predictive abilities obtained with G increased
markedly for all traits with the exception of fat\@.
The single-step combination of genomic and transriptomic information
($T^{*}|G$) yielded a small improvement over G alone for the traits dry matter
yield and protein.\@.
Compared to using only G, trait prediction did not benefit from the single-step
approach when T was not superior to G in the reduced set.
The only exception was protein.
Predictive abilities resulting from combinations of pedigree with genomic
($G^{*}|P$) information were higher than P alone for all traits and $T^{*}|P$
improved upon P in the `full' set for four traits.
Here, we did not consider the inclusion of dominance effects for the hybrid
material after a previous investigation into the predictive ability of various
``omics'' predictors in the same material concluded no considerable gains 
\cite{Westhues2017}.

\begin{figure}[H]
\centering
  \includegraphics{./tables_figures/pred_ability_hybrid.pdf}
  \caption{
    Predictive abilities and average, bootstrapped standard errors for the set
    of maize hybrids and seven agronomic traits.
    As predictors, pedigree (P), genomic (G), transcriptomic (T) data and 
    combinations thereof were used.
    The `Full' set of genotypes comprises 1,521 hybrid progeny based on 245
    parental inbred lines, whereas the `Reduced' set comprises 685 hybrid
    progeny based on 103 parental inbred lines.
  }
\label{fig:HybridResults}
\end{figure}

\Genetics2level{Impact of coverage by the incomplete predictor}
Predictive abilities for the six traits in the different core sets of Exp1
varied widely within most traits (Fig.~\ref{fig:core-results}).
For the majority of traits, an initial increase in predictive abilities was
observed when moving from a core set size of 10\% to 20\%/30\%, regardless of
whether G or T individually was the superior predictor.
For the traits cob weight, grain weight and plant height, which were predicted
with higher precision via G than via T in the reduced set of lines, predictive
abilities were higher when only a small number of genotypes was included in the
core set.
This corresponds to a situation where a large number of genotypes is being 
imputed and the majority of the information is derived from genomic information.
A hike in predictive abilities, when moving from a core fraction of 10\% to 
90\%, was observed for ear diameter and kernel width, which were predicted with
greater precision when using T individually compared to using G individually.
The only trait that was barely affected by a change in the core fraction was
days until silking for which the predictive abilities using either G or T
individually were almost identical (Fig.~\ref{fig:InbredResults}).

\begin{figure}[H]
  \centering
  \includegraphics{./tables_figures/pred_ability_inbred_core_fraction.pdf}
  \caption{
  Influence of the fraction of genotypes covered by the incomplete predictor
  on predictive abilities.
  Results are shown for the core set of 110 inbred lines from Exp1 for
  which both, G and T are available for any genotype and six agronomic traits.
  The fraction of genotypes (\textit{i.e.} `Core Fraction') for which
  T was available is given on the x-axis.
  Predictive abilities for core fractions between 0.1 and 0.9 are based on
  transcriptomic effects derived from a single-step prediction model involving
  transcriptomic (T) and genomic (G) data, respectively.
  The predictive ability at a core fraction of 0 is based only on G and at a
  core fraction of 1 it is based only on T.
  Dotted lines indicate that $T > G$, dashed lines indicate that $T \approx G$
  and solid lines indicate that $G > T$ when running predictions using single
  predictors on the same data set.
  }
\label{fig:core-results}
\end{figure}

\section*{Discussion}
\Genetics2level{Advantages of marker effect models}
Single-step approaches in hybrid populations have so far been limited to pig
breeding~\cite{Xiang2015,Xiang2016,Tusell2016}.
All of these studies used breeding value-based single-step BLUP methods
(SS-BLUP) conceived by \citeNP{Legarra2009} and \citeNP{Christensen2010} with
extensions to hybrid populations~\cite{Christensen2014,Christensen2015}.
An attractive property of SS-BLUP models is their computational advantage over 
marker-effect single-step Bayesian regression models (SSBR) when the number of 
genotypes is smaller than the number of features of the incomplete predictor.
Nevertheless, SS-BLUP suffers from a variety of issues.
First, the combination of the numerator with the genomic relationship matrix
requires commensurability of the two matrices, which necessitates some form of
scaling and weighting of $\mathbf{A}$ and $\mathbf{G}$
\cite{Christensen2012,Christensen2012a}.
Second, the fact that mostly recent individuals have genotypes whereas the 
majority of old individuals has only pedigree records introduces a bias that 
has to be accounted for~\cite{Vitezica2011,Legarra2015,Garcia-Baccino2017}.
Third, the addition of new genotypes requires updates of $\mathbf{G}$ that
demand additional measures such as computing the inverse of $\mathbf{G}$ via
recursion
\cite{Misztal2014,Misztal2016,Misztal2016a,Fragomeni2015,Masuda2016,Pocrnic2016}.
Advantages of the SSBR algorithm, conceived by \citeNP{Fernando2014}, include
that (i) it does not require commensurability of $\mathbf{A}$ and $\mathbf{G}$, 
(ii) it models the bias incurred by the incomplete predictor explicitly as 
$\mathbf{\epsilon}$ and (iii) its speed depends largely on the number of 
features.
To our best knowledge, this study presents the first implementation of the SSBR 
algorithm on a hybrid data set.

\Genetics2level{Addition of new predictors}
Pedigree records are the most ubiquitous predictor in animal and plant breeding
programs, because they are easy and inexpensive to collect.
Compared to pedigree information, which represent the expected relationship
among individuals, genomic information capture Mendelian sampling as well as the
effects of selection and thereby provide an improved proxy of the realized
relationship among individuals.
Nevertheless, genomic data do not exhaustively capture physiological epistasis 
\cite{Jiang2015,Guo2016,Vazquez2016}, describing the interactions 
within and between different biological strata~\cite{Sackton2016}.
Such interactions were found to be pervasive throughout the genomes of yeast 
\cite{Brem2005} and humans~\cite{Brown2014} and have motivated studies on the 
utility of downstream `omics' predictors for integrating such interactions.
Recently, encouraging results with regard to predicting complex traits using
`omics` data have been found for humans~\cite{Vazquez2016}, maize inbred lines
\cite{Guo2016} and maize hybrids~\cite{Westhues2017}.
While single-step prediction has become an established method when complete
pedigree but only incomplete genomic records are available, its application has
not yet been considered for the imputation of downstream `omics' predictors
such as transcriptomic, proteomic or metabolomic data.
High correlations between the off-diagonal elements of the kernels for G and T
and abundant interactions between distal SNPs and transcripts in hybrid maize
\cite{Westhues2017} suggested that the substitution of G with T as a single-step
predictor would be a logical next step.
Here, we considered single-step approaches for a maize inbred line diversity
panel as well as for a collection of maize hybrids using pedigree, genomic and
transcriptomic data as predictors.
While the repeatability of relationship coefficients and genomic markers can be
assumed to be equal to 1, this would not be the case for the aforementioned
quantitative `omics' predictors. 
This would negatively affect the precision of the imputation when using them as
imputors, which was not considered here.

\Genetics2level{Superiority of single-step models in line prediction}
When using a single-step approach based on complete genomic and incomplete
transcriptomic information, gains in predictive ability for the inbred lines
were at best just slightly higher than when using complete genomic information
alone.
One of the reasons might be that the subset of genotypes that was covered by
transcriptomic data was not a good representation of the full genetic space of
all available genotypes.
Another possibility is that the number of additional phenotypes in the full
set compared to the reduced set ($\Delta(n) = 54$) was not large enough to
provide considerably more useful information.
A larger dataset was available to~\citeN{Ashraf2016}, who imputed genotypes
for about 10,000 wheat lines using pedigree information and observed greater
prediction accuracy of the single-step method over genomic BLUP for all four
evaluated traits.
Predictive abilities reported for the maize diversity panel in this study were
slightly different from those reported by~\citeN{Guo2016}, for three
possible reasons:
First, we reduced the dataset to 164 tropical/subtropical lines with little
evident population structure and excluded all other genotypes from our analyses.
In contrast~\citeN{Guo2016} used genotypes from four different subpopulations
with a model using a fixed effect to account for the differences between
subpopulations.
Second, we applied quality checks for the predictor data after generating the
subset of 164 inbred lines, whereas~\citeN{Guo2016} applied the quality checks to
368 genotypes.
Third, while~\citeN{Guo2016} used five-fold cross-validation with 500 repetitions,
we employed a leave-one-out cross-validation scheme (LOOCV).
Notwithstanding, the relative differences between predictive abilities for G 
and T, respectively, were similar in both studies.

\Genetics2level{Superiority of single-step models in hybrid prediction}
Hybrid breeding is a particularly challenging application field for prediction
tools \cite{Kadam2016}.
Here, $2n$ parent individuals from two genetically distinct heterotic groups are
crossed to each other, yielding $n^{2}$ potential hybrid progeny that would
require intensive field testing.
In medium-sized plant breeding programs, the advent of the doubled haploid (DH) 
technology~\cite{Wedzony2009} now allows for an annual production of 1,000 
parent lines in each heterotic group, amounting to $10^{6}$ putative hybrid
progeny.
\citeN{Westhues2017} have shown that, in hybrid breeding, the probability 
of successfully selecting the best observed genotypes based on the best 
predicted candidates is a strongly convex function of the predictive ability.
Thus, even minor gains, when using transcriptomic and genomic data in a
single step, might justify additional investments in RNA-seq at least for a
subset of genotypes that represents the genetic space of the breeding 
material well.
For the hybrid data, the predictive abilities of G increased considerably in the
full set of genotypes compared to the reduced set for several agronomic traits.
Hence, the inclusion of transcriptomic data could not further improve upon
genomic information alone.
Predictive abilities obtained when using pedigree data could be 
improved considerably when, depending on the agronomic trait, combined with 
either transcriptomic or genomic information.
This suggests that single-step prediction should always be considered when the
incomplete predictor offers an appreciably higher predictive ability compared
to the complete predictor.
A previous study using the same data as in Exp2 concluded that modeling of
specific combining ability (SCA) effects did not appreciably improve predictive
abilities over that of a purely additive model for any of the investigated
traits~\cite{Westhues2017}.
Nevertheless, the inclusion of SCA effects in a single-step prediction model
would be as straightforward as computing the element-wise product between the
feature matrices $\mathbf{M}_{D}$ and $\mathbf{M}_{F}$ of the two corresponding
heterotic groups~\cite{Martini2016}.
While different studies have found that the low magnitude of SCA in hybrid maize
breeding programs typically does not warrant its inclusion, this situation
might be different in crops where heterotic patterns are not yet clearly
defined~\cite{Zhao2015}.

\Genetics2level{Coverage of the genetic space}
Whereas animal breeding populations are oftentimes so large that the assembly 
of a core set of individuals with records on multiple predictors can effectively 
be done at random once more than 10,000 animals have predictor information
\cite{Fragomeni2015,Lourenco2015,Masuda2016}, population sizes in plant breeding 
programs are typically considerably smaller.
Single-step prediction offers the promise of leveraging the predictive ability
for all genotypes by borrowing the superior performance of an auxiliary
predictor on only a subset of genotypes, thereby reducing predictor costs.
Hence, we were interested in determining to what extent individuals should be 
complemented with information on another predictor.
By randomly and repeatedly declaring transcriptomic information missing from
genotypes in Exp1 we examined how reliably single-step prediction works
depending on the fraction of genotypes covered by both, the complete and the
incomplete predictor.
We observed that about $x=20-30\%$ of genotypes need to be covered by both
predictors to match the performance achieved with the complete predictor
individually.
Beyond this fraction of covered genotypes the trend in predictive ability
changes was stable in either direction.
Predictive abilities increased simultaneously with $x$ when the incomplete
predictor outperformed the complete predictor in the reduced set of genotypes
and decreased with increasing $x$ when the complete predictor was the better
predictor.
This pattern clearly elucidates the influence of the difference between the
two predictors on predictive ability in a single-step approach.

\section*{Conclusions}
We successfully applied single-step prediction to inbred and hybrid data sets
while imputing with a quantitative downstream `omics' predictor.
By declaring different subsets of individuals as covered by one or two 
predictors, respectively, we could elucidate the influence of differences
between two predictors on predictive ability when using a single-step
approach.
When the discrepancy in predictive ability between the complete and the
incomplete predictor was skewed profoundly in favor of the incomplete predictor,
such as for pedigree and transcriptomic data, single-step prediction was shown
to be highly beneficial.
Extensions to more than two predictors were outside the scope of this study
but, with mounting interest in systems genetics, should be considered in the
future.

\section{Acknowledgments}
This project was funded by the German Federal Ministry of Education and 
Research (BMBF) within the projects OPTIMAL (FKZ: 0315958B,0315958F),
SYNBREED (FKZ: 0315528D) and by the German Research Foundation 
(DFG, Grants No. ME 2260/5-1 and SCHO 764/6-1).
The authors acknowledge support by the state of Baden-W{\"u}rttemberg through 
bwHPC\@.
Financial support for M.W. was provided by the Fiat Panis foundation, Ulm,
Germany.

\nolinenumbers
% Bibliography
\bibliography{library}
\bibliographystyle{mychicago}
\end{document}